---
title: "asian carp first part"
author: "Eddie Wu"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This .Rmd file is to show the progress on asian carp temperature and condition analyses. Since sub-sampling from spatial autocorrelation does not give significantly different results from normal analysis, we present the results without sub-sampling here.

For each species, we have four different models:

1. Simple linear model (same slope, same intercept)
2. Linear additive model (same slope, different intercept)
3. Interaction model (different slope, same intercept)
4. Group-specific model (different slope, different intercept)

And we consider two temperature metrics:

1. Annual temperature
2. Winter temperature (temperature from the coldest quarter)


```{r data import and setup}
library(ggplot2)
library(ggfortify)
library(dplyr)

## Import data
asian.carp <- read.csv("asian carp final.csv")
asian.carp$Condition <- as.factor(asian.carp$Condition)

Black <- read.csv("eddie_carp_new.csv")
Black$condition <- as.factor(Black$condition)

## Separate by species
Grass <- asian.carp[asian.carp$Species=="Grass",]
Bighead <- asian.carp[asian.carp$Species=="Bighead",]
Silver <- asian.carp[asian.carp$Species=="Silver",]
Big.sil <- rbind(Bighead, Silver) # combine the two groups

```

## Asian carp

### Using annual temperature
```{r asian carp annual temp}
# Clean data
asian.carp.clean <- asian.carp %>% 
  filter(Condition %in% c("natural", "artificial"))

# Build the models
asian.simple <- lm(log(AAM)~AnnualTemp, data = asian.carp.clean)
asian.linear <- lm(log(AAM)~AnnualTemp+Condition, data = asian.carp.clean)
asian.int <- lm(log(AAM)~AnnualTemp:Condition, data = asian.carp.clean)
asian.group <- lm(log(AAM)~AnnualTemp*Condition, data = asian.carp.clean)

# Compare the AICs
AIC(asian.simple, asian.linear, asian.int, asian.group)
```

For Asian carp using annual air temperature, the simple linear model works the best.


### Using cold temperature
```{r asian carp cold temp}
# Build the models
asian.simple <- lm(log(AAM)~ColdTemp, data = asian.carp.clean)
asian.linear <- lm(log(AAM)~ColdTemp+Condition, data = asian.carp.clean)
asian.int <- lm(log(AAM)~ColdTemp:Condition, data = asian.carp.clean)
asian.group <- lm(log(AAM)~ColdTemp*Condition, data = asian.carp.clean)

# Compare the AICs
AIC(asian.simple, asian.linear, asian.int, asian.group)
```

Same conclusion when using the cold temperature. But using annual temperature seems to have a better model as explained by lower AIC values.


## Grass carp

### Using annual temperature
```{r grass carp annual temp}
# Clean data
grass.clean <- Grass %>% 
  filter(Condition %in% c("natural", "artificial"))

# Build the models
grass.simple <- lm(log(AAM)~AnnualTemp, data = grass.clean)
grass.linear <- lm(log(AAM)~AnnualTemp+Condition, data = grass.clean)
grass.int <- lm(log(AAM)~AnnualTemp:Condition, data = grass.clean)
grass.group <- lm(log(AAM)~AnnualTemp*Condition, data = grass.clean)

# Compare the AICs
AIC(grass.simple, grass.linear, grass.int, grass.group)

```

For grass carp using annual air temperature, the simple linear model works the best. Although there does not seem to be much difference among the models.


### Using cold temperature
```{r grass carp cold temp}
# Build the models
grass.simple <- lm(log(AAM)~ColdTemp, data = grass.clean)
grass.linear <- lm(log(AAM)~ColdTemp+Condition, data = grass.clean)
grass.int <- lm(log(AAM)~ColdTemp:Condition, data = grass.clean)
grass.group <- lm(log(AAM)~ColdTemp*Condition, data = grass.clean)

# Compare the AICs
AIC(grass.simple, grass.linear, grass.int, grass.group)
```

Same as using annual temperature.


## Bighead and silver carp

### Using annual temperature
```{r bighead silver carp annual temp}
# Clean data
big.sil.clean <- Big.sil %>% 
  filter(Condition %in% c("natural", "artificial"))

# Build the models
big.sil.simple <- lm(log(AAM)~AnnualTemp, data = big.sil.clean)
big.sil.linear <- lm(log(AAM)~AnnualTemp+Condition, data = big.sil.clean)
big.sil.int <- lm(log(AAM)~AnnualTemp:Condition, data = big.sil.clean)
big.sil.group <- lm(log(AAM)~AnnualTemp*Condition, data = big.sil.clean)

# Compare the AICs
AIC(big.sil.simple, big.sil.linear, big.sil.int, big.sil.group)
```

For grass carp using annual air temperature, the simple linear model works the best. Although there does not seem to be much difference among the models.


### Using cold temperature
```{r bighead silver carp cold temp}
# Build the models
big.sil.simple <- lm(log(AAM)~ColdTemp, data = big.sil.clean)
big.sil.linear <- lm(log(AAM)~ColdTemp+Condition, data = big.sil.clean)
big.sil.int <- lm(log(AAM)~ColdTemp:Condition, data = big.sil.clean)
big.sil.group <- lm(log(AAM)~ColdTemp*Condition, data = big.sil.clean)

# Compare the AICs
AIC(big.sil.simple, big.sil.linear, big.sil.int, big.sil.group)
```

Same conclusion as annual temperature. Annual temperature fits better models.


## Black carp

### Using annual temperature
```{r black carp annual temp}
# Clean data
black.clean <- Black %>% filter(!row_number() == 15) %>% filter(sex != "male")

# Build the models
black.simple <- lm(log(AAM)~AnnualTemp, data = black.clean)
black.linear <- lm(log(AAM)~AnnualTemp+condition, data = black.clean)
black.int <- lm(log(AAM)~AnnualTemp:condition, data = black.clean)
black.group <- lm(log(AAM)~AnnualTemp*condition, data = black.clean)

# Compare the AICs
AIC(black.simple, black.linear, black.int, black.group)
```

The interaction model seems to perfrom slightly better than the others.

### Using cold temperature
```{r black carp cold temp}
# Build the models
black.simple <- lm(log(AAM)~ColdTemp, data = black.clean)
black.linear <- lm(log(AAM)~ColdTemp+condition, data = black.clean)
black.int <- lm(log(AAM)~ColdTemp:condition, data = black.clean)
black.group <- lm(log(AAM)~ColdTemp*condition, data = black.clean)

# Compare the AICs
AIC(black.simple, black.linear, black.int, black.group)
```

The linear additive model seems to be the best one, while the interaction model
is the worst fit.


## For black carp, if we remove the Southern Ukraine data point:

### Using annual temperature
``` {r black carp outlier removal annual}
# Remove the Southern Ukraine data point
black.clean.r <- black.clean[black.clean$AAM != 11,]

# Remove the Nucet farm data point
black.clean.r <- black.clean.r[black.clean$AAM != 4,]

# Build the models
black.simple.r <- lm(log(AAM)~AnnualTemp, data = black.clean.r)
black.linear.r <- lm(log(AAM)~AnnualTemp+condition, data = black.clean.r)
black.int.r <- lm(log(AAM)~AnnualTemp:condition, data = black.clean.r)
black.group.r <- lm(log(AAM)~AnnualTemp*condition, data = black.clean.r)

# Compare the AICs
AIC(black.simple.r, black.linear.r, black.int.r, black.group.r)
```

Linear addictive and interaction model.


### Using cold temperature
``` {r black carp outlier removal cold}
# Build the models
black.simple.r <- lm(log(AAM)~ColdTemp, data = black.clean.r)
black.linear.r <- lm(log(AAM)~ColdTemp+condition, data = black.clean.r)
black.int.r <- lm(log(AAM)~ColdTemp:condition, data = black.clean.r)
black.group.r <- lm(log(AAM)~ColdTemp*condition, data = black.clean.r)

# Compare the AICs
AIC(black.simple.r, black.linear.r, black.int.r, black.group.r)
```

Linear additive model.


## Concluding points

* For all the other asian carp species, the simple linear model seems to be the best fit. Besides, using cold temperature and annual temperature does not produce different results.

* For black carp, the linear additive model (*same slope, different intercept*) seems to be a better fit. Using cold temperature on the original data suggests that the interaction model is the best (probably because the winter temperature at these two locations are very similar.)

* After removing the two data points, we see that the linear additive still seems to be a better fit.


## Diagnostic plots

Only the diagnostic plots with cold temperature
``` {r d plots}
# Asian carp

autoplot(asian.simple)
autoplot(asian.linear)
autoplot(asian.int)
autoplot(asian.group)

# Grass carp
autoplot(grass.simple)
autoplot(grass.linear)
autoplot(grass.int)
autoplot(grass.group)

# Bighead and silver carp
autoplot(big.sil.simple)
autoplot(big.sil.linear)
autoplot(big.sil.int)
autoplot(big.sil.group)

# Black carp
autoplot(black.simple)
autoplot(black.linear)
autoplot(black.int)
autoplot(black.group)

# Black carp modified
autoplot(black.simple.r)
autoplot(black.linear.r)
autoplot(black.int.r)
autoplot(black.group.r)

```
