---
title: "Spatial autocorrelation"
author: "Eddie Wu"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This .Rmd file is to test the spatial autocorrelation in our black carp data file, and try to find the best distance that reduces the spatial autocorrelation when doing subsampling. In this file, we:

1. Show the distribution of our data points on a map.

1. Use moran's I test to test the global spatial autocorrelation in our dataset.

2. Use the correlog plots to examine how spatial autocorrelation changes with distances.

3. Use the variogram to further test point 2.

4. Subsample at 250 km and 550 km, and check the local moran's I after sub-sampling to see if spatial autocorrelation has been reduced.

5. Conduct latitudinal stratification analysis.

6. Look at the Chinese dataset - 17 data points.



```{r library, warning=FALSE, message=FALSE}
library(gstat)
library(ggplot2)
library(dplyr)
library(spdep)
library(sp)
library(nlme)
library(ape)
library(MuMIn)
library(raster)
library(ncf)
library(knitr)
library(rnaturalearth)
library(sf)
```

```{r data import}
## Import location data
location <- read.csv("location_no_temps.csv")
location <- unique(location)

location.new <- location %>% # remove duplicating locations
  filter(Location != "South Ukraine") # remove South Ukraine

# Clean the data
carp <- read.csv("eddie_carp_new.csv")
carp.r <- carp %>% 
  filter(!row_number() == 5) %>%
  filter(sex != "male")


## Download one file to get the spatial points
# (this defines what projection to use when converting to spatial objects)
tmin.1979 <- brick("cpc/tmin.1979.nc", varname = "tmin")
tmin.1979<-rotate(tmin.1979)


## Define a function to plot the correlogram for local moran's I

plot_correlog <- function(model){
  # Define margins
  par(mfrow=c(2,1),mar=c(4,4,2,2))

  # Use the coorelog function to develop the relationship
  test <- correlog(coords$long, coords$lat, model$residuals,
                   increment=50, resamp=500, latlon=T)

  # Plot with the entire distance range
  plot(test, main=paste(model$call))
  abline(h=0)

  # Reduce the distance range to 2500 km
  plot(test, main="", xlim=c(0,2500))
  abline(h=0)
}
```


## Data distribution

```{r distribution}
## Get the world data
world <- ne_countries(scale = "small", returnclass = "sf")

## Plots (global plot)
world %>% 
  filter(admin != "Antarctica") %>%
  # remove antarctica
  ggplot()+
  geom_sf()+
  geom_point(aes(x=Longitude, y=Latitude), data = location,
             size = 1, color = "darkred")+
  theme_bw()+
  xlab("Longitude") + ylab("Latitude")
```


## Section 1: Spatial autocorrelation examination for entire dataset

### Model definition

We first need to define all 12 models (3 models for each temperature).

```{r model with SU}
# Define the models for AnnualTemp
annual.lin <- lm(log(AAM)~AnnualTemp, data = carp.r)
annual.add <- lm(log(AAM)~AnnualTemp+condition, data = carp.r)
annual.int <- lm(log(AAM)~AnnualTemp*condition, data = carp.r)

# Define the models for ColdTemp
cold.lin <- lm(log(AAM)~ColdTemp, data = carp.r)
cold.add <- lm(log(AAM)~ColdTemp+condition, data = carp.r)
cold.int <- lm(log(AAM)~ColdTemp*condition, data = carp.r)

# Define the model for WarmTemp
warm.lin <- lm(log(AAM)~WarmTemp, data = carp.r)
warm.add <- lm(log(AAM)~WarmTemp+condition, data = carp.r)
warm.int <- lm(log(AAM)~WarmTemp*condition, data = carp.r)

# Define the models for GDD0
gdd.lin <- lm(log(AAM)~average_gdd_0, data = carp.r)
gdd.add <- lm(log(AAM)~average_gdd_0+condition, data = carp.r)
gdd.int <- lm(log(AAM)~average_gdd_0*condition, data = carp.r)
```


### Global Moran.I

```{r global}
## Make spatial dataframe
coords <- data.frame("long"=location[,3],"lat"=location[,2])
df <- data.frame(a = 1:nrow(location[3]))
spatial.data <- SpatialPointsDataFrame(coords,df,proj4string = tmin.1979@crs)

# Get a distance matrix from all points
dists <- spDists(spatial.data, longlat = TRUE)


## Global Moran.I
# Annual Temp
Moran.annual.lin <- Moran.I(annual.lin$residuals, dists)
Moran.annual.add <- Moran.I(annual.add$residuals, dists)
Moran.annual.int <- Moran.I(annual.int$residuals, dists)

global.moran.annual <- data.frame(
  Model = c("simple linear", "linear additive", "interaction"),
  Observed = c(Moran.annual.lin$observed, Moran.annual.add$observed,
               Moran.annual.int$observed),
  Expected = c(Moran.annual.lin$expected, Moran.annual.add$expected,
               Moran.annual.int$expected),
  sd = c(Moran.annual.lin$sd, Moran.annual.add$sd, Moran.annual.int$sd),
  p.value = c(Moran.annual.lin$p.value, Moran.annual.add$p.value,
              Moran.annual.int$p.value)
)

kable(global.moran.annual)

# Cold Temp
Moran.cold.lin <- Moran.I(cold.lin$residuals, dists)
Moran.cold.add <- Moran.I(cold.add$residuals, dists)
Moran.cold.int <- Moran.I(cold.int$residuals, dists)

global.moran.cold <- data.frame(
  Model = c("simple linear", "linear additive", "interaction"),
  Observed = c(Moran.cold.lin$observed, Moran.cold.add$observed,
               Moran.cold.int$observed),
  Expected = c(Moran.cold.lin$expected, Moran.cold.add$expected,
               Moran.cold.int$expected),
  sd = c(Moran.cold.lin$sd, Moran.cold.add$sd, Moran.cold.int$sd),
  p.value = c(Moran.cold.lin$p.value, Moran.cold.add$p.value,
              Moran.cold.int$p.value)
)

kable(global.moran.cold)

# Warm Temp
Moran.warm.lin <- Moran.I(warm.lin$residuals, dists)
Moran.warm.add <- Moran.I(warm.add$residuals, dists)
Moran.warm.int <- Moran.I(warm.int$residuals, dists)

global.moran.warm <- data.frame(
  Model = c("simple linear", "linear additive", "interaction"),
  Observed = c(Moran.warm.lin$observed, Moran.warm.add$observed,
               Moran.warm.int$observed),
  Expected = c(Moran.warm.lin$expected, Moran.warm.add$expected,
               Moran.warm.int$expected),
  sd = c(Moran.warm.lin$sd, Moran.warm.add$sd, Moran.warm.int$sd),
  p.value = c(Moran.warm.lin$p.value, Moran.warm.add$p.value,
              Moran.warm.int$p.value)
)

kable(global.moran.warm)

# GDD0
Moran.gdd.lin <- Moran.I(gdd.lin$residuals, dists)
Moran.gdd.add <- Moran.I(gdd.add$residuals, dists)
Moran.gdd.int <- Moran.I(gdd.int$residuals, dists)

global.moran.gdd <- data.frame(
  Model = c("simple linear", "linear additive", "interaction"),
  Observed = c(Moran.gdd.lin$observed, Moran.gdd.add$observed,
               Moran.gdd.int$observed),
  Expected = c(Moran.gdd.lin$expected, Moran.gdd.add$expected,
               Moran.warm.int$expected),
  sd = c(Moran.gdd.lin$sd, Moran.gdd.add$sd, Moran.gdd.int$sd),
  p.value = c(Moran.gdd.lin$p.value, Moran.gdd.add$p.value,
              Moran.gdd.int$p.value)
)

kable(global.moran.gdd)
```

There is no global spatial autocorrelation on the entire dataset for all three temperature metrics.


### Local spatial autocorrelation

```{r correlog}
## Annual temperature
plot_correlog(annual.lin)
plot_correlog(annual.add)
plot_correlog(annual.int)

## Cold temperature
plot_correlog(cold.lin)
plot_correlog(cold.add)
plot_correlog(cold.int)

## Warm temperature
plot_correlog(warm.lin)
plot_correlog(warm.add)
plot_correlog(warm.int)

## GDD0
plot_correlog(gdd.lin)
plot_correlog(gdd.add)
plot_correlog(gdd.int)
```

In general, the correlog plots suggest that for both temperature, 550 km is a good distance to reduce local spatial autocorrelation.

If we pick 250 km as the cutoff distance for sub-sampling, we would have 20 location points.

If we pick 550 km as the cutoff distance for sub-sampling, we would have only 12 location points.


## Section 2: Subsampling

### Subsampling at 250km

Here, we try to sub-sample at 250km and check the local moran's I (correlog plot) after sub-sampling.

#### Local moran's I - annual 250

```{r subsample local moran annual 250, fig.width=14, fig.height=12, results='hide'}
# Store the r2 values
r2.250.raw <- matrix(NA,10, 2)
colnames(r2.250.raw) <- c("annual","cold")

table(carp.r$spatial.code.250)

par(mfrow = c(5, 2))
par(mar=c(1,1,1,1))

## Check the local moran's results after sub-sampling
for (i in 1:10){
  sub <- carp.r %>% group_by(spatial.code.250) %>% sample_n(size=1)
  reg.sub.annual <- lm(log(sub$AAM)~sub$AnnualTemp)
  test <- correlog(sub$longitude, sub$latitude, reg.sub.annual$residuals,
                   increment=50, resamp=500, latlon=T)
  plot(test, main="", xlim=c(0,2500))
  abline(h=0)
  r2.250.raw[i,1] <- summary(reg.sub.annual)$adj.r.squared #get the r2
}
```


#### Local moran's I - cold 250

```{r subsample local moran cold 250, fig.width=14, fig.height=12, results='hide'}
table(carp.r$spatial.code.250)

par(mfrow = c(5, 2))
par(mar=c(1,1,1,1))

## Check the local moran's results after sub-sampling
for (i in 1:10){
  sub <- carp.r %>% group_by(spatial.code.250) %>% sample_n(size=1)
  reg.sub.cold <- lm(log(sub$AAM)~sub$ColdTemp)
  test <- correlog(sub$longitude, sub$latitude, reg.sub.cold$residuals,
                   increment=50, resamp=500, latlon=T)
  plot(test, main="", xlim=c(0,2500))
  abline(h=0)
  r2.250.raw[i,2] <- summary(reg.sub.cold)$adj.r.squared #get the r2
}
```


Subsampling at 250 km does not reduce the spatial autocorrelation in our dataset.


### Subsampling at 550km

Now, we try to subsample at a larger distance - 550km.

#### Local moran's I - annual 550

```{r subsample local moran annual 550, fig.width=14, fig.height=12, results='hide'}
table(carp.r$spatial.code.550)

# Store the r2 values
r2.550.raw <- matrix(NA,10, 2)
colnames(r2.550.raw) <- c("annual","cold")

par(mfrow = c(5, 2))
par(mar=c(1,1,1,1))

## Check the local moran's results after sub-sampling
for (i in 1:10){
  sub <- carp.r %>% group_by(spatial.code.550) %>% sample_n(size=1)
  reg.sub.annual <- lm(log(sub$AAM)~sub$AnnualTemp)
  test <- correlog(sub$longitude, sub$latitude, reg.sub.annual$residuals,
                   increment=50, resamp=500, latlon=T)
  plot(test, main="", xlim=c(0,2500))
  abline(h=0)
  r2.550.raw[i,1] <- summary(reg.sub.annual)$adj.r.squared #get the r2
}
```


#### Local moran's I - cold 550

```{r subsample local moran cold 550, fig.width=14, fig.height=12, results='hide'}
table(carp.r$spatial.code.550)

par(mfrow = c(5,2))
par(mar=c(1,1,1,1))

## Check the local moran's results after sub-sampling
for (i in 1:10){
  sub <- carp.r %>% group_by(spatial.code.550) %>% sample_n(size=1)
  reg.sub.cold <- lm(log(sub$AAM)~sub$ColdTemp)
  test <- correlog(sub$longitude, sub$latitude, reg.sub.cold$residuals,
                   increment=50, resamp=500, latlon=T)
  plot(test, main="", xlim=c(0,2500))
  abline(h=0)
  r2.550.raw[i,2] <- summary(reg.sub.cold)$adj.r.squared #get the r2
}
```

We have seen that subsampling at 550 km would largely reduce the spatial autocorrelation in our dataset at low distances, even though we are only left with 12 data points.


### Compare the r2 values

```{r compare r2}
r2 <- data.frame(
  Time = c("Original", "Sub 250","Sub 550"),
  Annual = c(summary(annual.lin)$adj.r.squared,
             mean(unique(r2.250.raw[,1])),
             mean(unique(r2.550.raw[,1]))),
  Cold = c(summary(cold.lin)$adj.r.squared,
           mean(unique(r2.250.raw[,2])),
           mean(unique(r2.550.raw[,2])))
)
kable(r2)
```

However, subsampling at 550 km would reduce our R2 value as well, suggesting that the model would not fit for the sub-dataset.



### Latitudinal stratification

Now we would like to explore if stratified sub-sampling using latitude would reduce our spatial autocorrelation at low distances.

#### one degree latitude bin - annual

```{r latitude one annual, fig.width=14, fig.height=12, results='hide'}

table(carp.r$lat.one)

par(mfrow = c(5,2))
par(mar=c(1,1,1,1))

for (i in 1:10){
  sub <- carp.r %>% group_by(lat.one) %>% sample_n(size=1)
  reg.sub.annual <- lm(log(sub$AAM)~sub$AnnualTemp)
  test <- correlog(sub$longitude, sub$latitude, reg.sub.annual$residuals,
                   increment=50, resamp=500, latlon=T)
  plot(test, main="", xlim=c(0,2500))
  abline(h=0)
}

dev.off()
```

#### one degree latitude bin - cold

```{r latitude one cold, fig.width=14, fig.height=12, results='hide'}

table(carp.r$lat.one)

par(mfrow = c(5,2))
par(mar=c(1,1,1,1))

for (i in 1:10){
  sub <- carp.r %>% group_by(lat.one) %>% sample_n(size=1)
  reg.sub.cold <- lm(log(sub$AAM)~sub$ColdTemp)
  test <- correlog(sub$longitude, sub$latitude, reg.sub.cold$residuals,
                   increment=50, resamp=500, latlon=T)
  plot(test, main="", xlim=c(0,2500))
  abline(h=0)
}

dev.off()
```

* For some subsamples, the negative autocorrelation at 550 km is reduced.

* It seems that the South Ukarine data point is the main cause that we have negative spatial autocorrelation at around 550 km. Therefore, we ought to remove this datapoint from our analyses.



## Section 3: Spatial autocorrelation without SU

Now we would like to re-examine the global and local spatial autocorrelation with the South Ukraine data point removed.


### Model construction

Again, we first need to define all 9 models (3 models for each temperature). This time we remove the SU data point.

```{r model SU removal}
# Remove the SU data point
carp.new <- carp.r %>% 
  filter(!row_number() == 20) # remove South Ukraine

# Define the models for AnnualTemp
annual.lin <- lm(log(AAM)~AnnualTemp, data = carp.new)
annual.add <- lm(log(AAM)~AnnualTemp+condition, data = carp.new)
annual.int <- lm(log(AAM)~AnnualTemp*condition, data = carp.new)

# Define the models for ColdTemp
cold.lin <- lm(log(AAM)~ColdTemp, data = carp.new)
cold.add <- lm(log(AAM)~ColdTemp+condition, data = carp.new)
cold.int <- lm(log(AAM)~ColdTemp*condition, data = carp.new)

# Define the model for WarmTemp
warm.lin <- lm(log(AAM)~WarmTemp, data = carp.new)
warm.add <- lm(log(AAM)~WarmTemp+condition, data = carp.new)
warm.int <- lm(log(AAM)~WarmTemp*condition, data = carp.new)

# Define the models for GDD0
gdd.lin <- lm(log(AAM)~average_gdd_0, data = carp.new)
gdd.add <- lm(log(AAM)~average_gdd_0+condition, data = carp.new)
gdd.int <- lm(log(AAM)~average_gdd_0*condition, data = carp.new)
```

### Global Moran.I

```{r global SU removal}
## Make spatial dataframe
coords <- data.frame("long"=location.new[,3],"lat"=location.new[,2])
df <- data.frame(a = 1:nrow(location.new[3]))
spatial.data <- SpatialPointsDataFrame(coords,df,proj4string = tmin.1979@crs)

# Get a distance matrix from all points
dists <- spDists(spatial.data, longlat = TRUE)


## Global Moran.I
# Annual Temp
Moran.annual.lin <- Moran.I(annual.lin$residuals, dists)
Moran.annual.add <- Moran.I(annual.add$residuals, dists)
Moran.annual.int <- Moran.I(annual.int$residuals, dists)

global.moran.annual <- data.frame(
  Model = c("simple linear", "linear additive", "interaction"),
  Observed = c(Moran.annual.lin$observed, Moran.annual.add$observed,
               Moran.annual.int$observed),
  Expected = c(Moran.annual.lin$expected, Moran.annual.add$expected,
               Moran.annual.int$expected),
  sd = c(Moran.annual.lin$sd, Moran.annual.add$sd, Moran.annual.int$sd),
  p.value = c(Moran.annual.lin$p.value, Moran.annual.add$p.value,
              Moran.annual.int$p.value)
)

kable(global.moran.annual)

# Cold Temp
Moran.cold.lin <- Moran.I(cold.lin$residuals, dists)
Moran.cold.add <- Moran.I(cold.add$residuals, dists)
Moran.cold.int <- Moran.I(cold.int$residuals, dists)

global.moran.cold <- data.frame(
  Model = c("simple linear", "linear additive", "interaction"),
  Observed = c(Moran.cold.lin$observed, Moran.cold.add$observed,
               Moran.cold.int$observed),
  Expected = c(Moran.cold.lin$expected, Moran.cold.add$expected,
               Moran.cold.int$expected),
  sd = c(Moran.cold.lin$sd, Moran.cold.add$sd, Moran.cold.int$sd),
  p.value = c(Moran.cold.lin$p.value, Moran.cold.add$p.value,
              Moran.cold.int$p.value)
)

kable(global.moran.cold)

# Warm Temp
Moran.warm.lin <- Moran.I(warm.lin$residuals, dists)
Moran.warm.add <- Moran.I(warm.add$residuals, dists)
Moran.warm.int <- Moran.I(warm.int$residuals, dists)

global.moran.warm <- data.frame(
  Model = c("simple linear", "linear additive", "interaction"),
  Observed = c(Moran.warm.lin$observed, Moran.warm.add$observed,
               Moran.warm.int$observed),
  Expected = c(Moran.warm.lin$expected, Moran.warm.add$expected,
               Moran.warm.int$expected),
  sd = c(Moran.warm.lin$sd, Moran.warm.add$sd, Moran.warm.int$sd),
  p.value = c(Moran.warm.lin$p.value, Moran.warm.add$p.value,
              Moran.warm.int$p.value)
)

kable(global.moran.warm)

# GDD0
Moran.gdd.lin <- Moran.I(gdd.lin$residuals, dists)
Moran.gdd.add <- Moran.I(gdd.add$residuals, dists)
Moran.gdd.int <- Moran.I(gdd.int$residuals, dists)

global.moran.gdd <- data.frame(
  Model = c("simple linear", "linear additive", "interaction"),
  Observed = c(Moran.gdd.lin$observed, Moran.gdd.add$observed,
               Moran.gdd.int$observed),
  Expected = c(Moran.gdd.lin$expected, Moran.gdd.add$expected,
               Moran.warm.int$expected),
  sd = c(Moran.gdd.lin$sd, Moran.gdd.add$sd, Moran.gdd.int$sd),
  p.value = c(Moran.gdd.lin$p.value, Moran.gdd.add$p.value,
              Moran.gdd.int$p.value)
)

kable(global.moran.gdd)
```

There is no global spatial autocorrelation on the entire dataset for all three temperature metrics.


### Local spatial autocorrelation

```{r correlog SU removal}
## Annual temperature
plot_correlog(annual.lin)
plot_correlog(annual.add)
plot_correlog(annual.int)

## Cold temperature
plot_correlog(cold.lin)
plot_correlog(cold.add)
plot_correlog(cold.int)

## Warm temperature
plot_correlog(warm.lin)
plot_correlog(warm.add)
plot_correlog(warm.int)

## GDD0
plot_correlog(gdd.lin)
plot_correlog(gdd.add)
plot_correlog(gdd.int)
```


## Section 4: China Dataset analyses

Now let's look at the datapoints in China. There are 17 datapoints.

```{r china distribution}
## Subset the Chinese dataset
black.china <- carp.r[carp.r$china == "y",]

## Get the world data
world <- ne_countries(scale = "small", returnclass = "sf")

## Plots (china plot)
world %>% 
  filter(admin == "China" | admin =="Taiwan") %>%
  # select only China
  ggplot()+
  geom_sf()+
  geom_point(aes(x=longitude, y=latitude), data = black.china,
             size = 2, color = "darkred")+
  theme_bw()+
  xlab("Longitude") + ylab("Latitude")
```

Note that the datapoint outside of the Chinese region is "Northern China". It is calculated through geographical center.

### Global moran's I

```{r global china}
# Run the models
lm.annual.china <- lm(log(AAM)~AnnualTemp, data = black.china)
summary(lm.annual.china)
lm.cold.china <- lm(log(AAM)~ColdTemp, data = black.china)
summary(lm.cold.china)

## Make spatial dataframe
coords <- data.frame("long"=black.china[,14],"lat"=black.china[,13])
df <- data.frame(a = 1:nrow(black.china[14]))
spatial.data <- SpatialPointsDataFrame(coords,df,proj4string = tmin.1979@crs)

# Get a distance matrix from all points
dists <- spDists(spatial.data, longlat = TRUE)

## Run the Moran.I test on the residuals
Moran.annual <- Moran.I(lm.annual.china$residuals, dists)
Moran.cold <- Moran.I(lm.cold.china$residuals, dists)

global.moran <- data.frame(
  Model = c("Moran.annual", "Moran.cold"),
  Observed = c(Moran.annual$observed, Moran.cold$observed),
  Expected = c(Moran.annual$expected, Moran.cold$expected),
  sd = c(Moran.annual$sd, Moran.cold$sd),
  p.value = c(Moran.annual$p.value, Moran.cold$p.value)
)

kable(global.moran)
```

No global spatial autocorrelation in the Chinese dataset.


### Local moran's I - correlog

```{r correlog china}
## Annual temperature
par(mfrow=c(2,1),mar=c(4,4,2,2))

# Use the coorelog function to develop the relationship
test <- correlog(black.china$longitude,
                 black.china$latitude, lm.annual.china$residuals,
                 increment=50, resamp=500, latlon=T)

# Plot with the entire distance range
plot(test,
     main="Annual Average Temperature Regression Residuals for Chinese data")
abline(h=0)
# Reduce the distance range to 1500 km
plot(test, main="", xlim=c(0,1500))
abline(h=0)


## Cold temperature
par(mfrow=c(2,1),mar=c(4,4,2,2))
# Use the coorelog function to develop the relationship
test <- correlog(black.china$longitude, 
                 black.china$latitude, lm.cold.china$residuals,
                 increment=50, resamp=500, latlon=T)

# Plot with the entire distance range
plot(test,
     main="Cold Quarter Temperature Regression Residuals for Chinese data")
abline(h=0)

# Reduce the distance range to 1500 km
plot(test, main="", xlim=c(0,1500))
abline(h=0)
```


### Subsample at 250km for Chinese dataset

Now we subsample at 250km to reduce spatial autocorrelation.

```{r china sub 250, fig.width=14, fig.height=12, results='hide'}
table(black.china$spatial.code.250)

## Annual temperature
par(mfrow=c(5,2), mar=c(1,1,2,2))
for (i in 1:10){
  sub <- black.china %>% group_by(spatial.code.250) %>% sample_n(size=1)
  reg.sub.annual <- lm(log(sub$AAM)~sub$AnnualTemp)
  
  test <- correlog(sub$longitude, sub$latitude, reg.sub.annual$residuals,
                   increment=50, resamp=500, latlon=T)
  plot(test, main="", xlim=c(0,1500))
  abline(h=0)
}

## Cold temperature
par(mfrow=c(5,2), mar=c(1,1,2,2))
for (i in 1:10){
  sub <- black.china %>% group_by(spatial.code.250) %>% sample_n(size=1)
  reg.sub.cold <- lm(log(sub$AAM)~sub$ColdTemp)
  
  test <- correlog(sub$longitude, sub$latitude, reg.sub.cold$residuals,
                   increment=50, resamp=500, latlon=T)
  plot(test, main="", xlim=c(0,1500))
  abline(h=0)
}
```



## Conclusions

* There was no global spatial autocorrelation in the entire dataset, but when looking at local clustering, we saw a strong negative autocorrelation at around 550 km.

* We tried to subsample at 250 km, and 550 km. 220 km did not reduce the spatial autocorrelation we saw. 550 km did reduce that, but left us with too few data to fit a model.

* We then tried to subsample by latitudinal straitification. We found that the negative autocorrelation was reduced in some iterations (when South Ukarine was not selected). We hypothesized that particular datapoint is the cause.

* We removed the SU data point and ran the local moran's I again. This time, we have no local clustering at 550 km!
