---
title: "Predicting age at maturity in black carp"
author: "Eddie Wu"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

We used different temperature metrics to predict black carp age at maturity.
Simple linear regression model was used for all metrics.


1. Annual average temperature.
**(obtained from the WORLDCLIM dataset BIO1)**
**(1970 - 2000)**

2. Cold quarter temperature.
**(obtained from the WORLDCLIM dataset BIO11)**
**(1970 - 2000)**

3. Warm quarter temperature.
**(obtained from the WORLDCLIM dataset BIO10)**
**(1970 - 2000)**

4. Growing degree day Base 0.
**(calculated from CPC Global Daily Temperature)**
**(1979 - 2022)**

5. Winter duration
**(calculated from CPC Global Daily Temperature)**
**(1979 - 2022)**

6. Annual water temperature.
**(averaged weekly water temperature calculated from futureStream database)**
**(1979 - 2005)**

7. Cold quarter water temperature.
**(averaged weekly water temperature calculated from futureStream database)**
**(1979 - 2005)**

Both water temperature metrics using the calculation method in the WorldClim database in terms of "annual" and "cold quarter". (package dismo)


## Predict black carp age at maturity
```{r data import, message=FALSE, warning=FALSE}
library(car)
library(dplyr)

## Import data
Black <- read.csv("black_carp_final.csv")
Black$condition <- as.factor(Black$condition)

# Clean data
Black <- Black %>% filter(!row_number() == 5) %>% filter(sex != "male")

# Remove the South Ukraine data point
black.clean <- Black %>% filter(!row_number() == 20)
```

```{r AAM prediction}
black.annual <- lm(log(AAM)~AnnualTemp, data = black.clean)
black.cold <- lm(log(AAM)~ColdTemp, data = black.clean)
black.warm <- lm(log(AAM)~WarmTemp, data = black.clean)
black.gdd <- lm(log(AAM)~average_gdd_0, data = black.clean)
black.below5 <- lm(log(AAM)~below5, data = black.clean)
black.water <- lm(log(AAM)~WaterTemp, data = black.clean)
black.coldwater <- lm(log(AAM)~WaterCold, data = black.clean)


summary(black.annual)
summary(black.cold)
summary(black.warm)
summary(black.gdd)
summary(black.below5)
summary(black.water)
summary(black.coldwater)


plot(black.annual, which=5)
```
