---
title: "Black/Asian carp model selection"
author: "Eddie Wu"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This .Rmd file is to show the progress on black carp and asian carp temperature and condition analyses. Since sub-sampling from spatial autocorrelation does not give significantly different results from normal analysis for black carp, we present the results without sub-sampling here. For other Asian carp species, we still sub-sample at a distance of 250 km.

SECTION 1 - Black Carp:

1. We first check how condition affects the relationship between black carp age at maturity and temperature by looking at three candidate models, for all six temperatures (annual, cold, warm, gdd0, water annual, water cold):

* Simple linear model (same slope, same intercept)
* Linear additive model (same slope, different intercept)
* Interaction model (different slope, different intercept)

2. Since we found that none of the interaction terms is significant, we will use only the simple linear model for prediction.

3. Since there is no significant relationship between warm temperature and age at maturity, we remove in from the following analyses.


SECTION 2 - Asian carp:

1. We conducted the same analyses on other asian carp species.


```{r data import, message=FALSE, warning=FALSE}
library(ggplot2)
library(ggfortify)
library(dplyr)
library(knitr)
library(tidyverse)
library(AICcmodavg) # for AICc and akaike weights
library(pwr)

## Import data
asian.carp <- read.csv("asian carp final.csv")
asian.carp$Condition <- as.factor(asian.carp$Condition)

Black <- read.csv("eddie_carp_new.csv")
Black$condition <- as.factor(Black$condition)

## Separate by species
Grass <- asian.carp[asian.carp$Species=="Grass",]
Bighead <- asian.carp[asian.carp$Species=="Bighead",]
Silver <- asian.carp[asian.carp$Species=="Silver",]
Big.sil <- rbind(Bighead, Silver) # combine the two groups


## Define two functions for AICs
compute_akaike_weights <- function(aic_scores) {
  # Find the AIC of the best model
  aic_min <- min(aic_scores)
  
  # Calculate delta AIC values
  d_aic <- aic_scores - aic_min
  
  # Compute Akaike weights
  akaike_weights <- exp(-0.5 * d_aic) / sum(exp(-0.5 * d_aic))
  
  return(akaike_weights)
}
compare_aic_scores <- function(aic_scores) {
  # Find the AIC of the best model
  aic_min <- min(aic_scores)
  
  # Determining if the smallest value is 2 units smaller than the others
  is_smaller_by_two <- all(aic_min + 2 <= aic_scores[aic_scores != aic_min])
  
  # Return the index if 
  if (is_smaller_by_two) {
    min_index <- which(aic_scores == aic_min)
    return(min_index)
  } else {
    return(-999)
  }
}
```



## SECTION 1: Black carp

For black carp data, we do not subsample at any distances. But we removed the South Ukraine data point for all the following analyses.


### Data cleaning

```{r data cleaning}
# Clean data
Black <- Black %>% filter(!row_number() == 5) %>% filter(sex != "male")

# Remove the South Ukraine data point
black.clean <- Black %>% filter(!row_number() == 20)
```


### Check slopes for three models (all six temperatures)
#### Annual Temperature
```{r black carp annual temp}
# Build the models
black.simple <- lm(log(AAM)~AnnualTemp, data = black.clean)
black.int <- lm(log(AAM)~AnnualTemp*condition, data = black.clean)


## Look at the summary (especially the slope for each model)
summary(black.simple)
summary(black.int)
```

* There is no significant interaction term or additive term in black carp using **annual temperature**, thus the simple regression model is the best.


#### Cold temperature
```{r black carp cold temp}
# Build the models
black.simple <- lm(log(AAM)~ColdTemp, data = black.clean)
black.int <- lm(log(AAM)~ColdTemp*condition, data = black.clean)


## Look at the summary (especially the slope for each model)
summary(black.simple)
summary(black.int)
```

* There is no significant interaction term or additive term in black carp using cold temperature, thus the simple regression model is the best.


#### Warm temperature
```{r black carp warm temp}
# Build the models
black.simple <- lm(log(AAM)~WarmTemp, data = black.clean)
black.int <- lm(log(AAM)~WarmTemp*condition, data = black.clean)


## Look at the summary (especially the slope for each model)
summary(black.simple)
summary(black.int)
```

* Warm temperature is not a significant predictor of black carp age at maturity (same conclusion as Madison found on other Asian carps), thus we remove it in the following analyses.


#### Base 0 annual growing degree day
```{r black carp gdd0}
# Build the models
black.simple <- lm(log(AAM)~average_gdd_0, data = black.clean)
black.int <- lm(log(AAM)~average_gdd_0*condition, data = black.clean)


## Look at the summary (especially the slope for each model)
summary(black.simple)
summary(black.int)
```

* There is no significant interaction term or additive term in black carp ggd0, thus the simple regression model is the best.


#### Annual water temperature
```{r black carp water annual}
# Build the models
black.simple <- lm(log(AAM)~WaterTemp, data = black.clean)
black.int <- lm(log(AAM)~WaterTemp*condition, data = black.clean)


## Look at the summary (especially the slope for each model)
summary(black.simple)
summary(black.int)
```

* There is no significant interaction term or additive term in black carp annual water temperature, thus the simple regression model is the best.


#### Cold water temperature
```{r black carp water cold}
# Build the models
black.simple <- lm(log(AAM)~WaterCold, data = black.clean)
black.int <- lm(log(AAM)~WaterCold*condition, data = black.clean)


## Look at the summary (especially the slope for each model)
summary(black.simple)
summary(black.int)
```

* There is a significant interaction term in black carp using cold quarter water temperature, thus we need to further decide based on R2 and AICc values.


#### Winter duration
```{r black carp winter duration}
# Build the models
black.simple <- lm(log(AAM)~below5, data = black.clean)
black.int <- lm(log(AAM)~below5*condition, data = black.clean)


## Look at the summary (especially the slope for each model)
summary(black.simple)
summary(black.int)
```

* There is no significant interaction term or additive term in black carp winter duration, thus the simple regression model is the best.


### Model selection for cold water temperature (sig interaction)
```{r black carp water cold model selection}
# Build the models
black.simple <- lm(log(AAM)~WaterCold, data = black.clean)
black.linear <- lm(log(AAM)~WaterCold+condition, data = black.clean)
black.int <- lm(log(AAM)~WaterCold*condition, data = black.clean)

## Get a table of corrected AICs and their Akaike weights
models <- list(black.simple, black.linear, black.int)
mod.names <- c('simple linear', 'linear additive', 'interaction')
aictab(cand.set = models, modnames = mod.names, sort = FALSE)


## R^2 value for three models
r_2 <- data.frame(
  Model = c("Simple linear", "Linear additive", "Interaction"),
  R2 = c(summary(black.simple)$adj.r.squared,
         summary(black.linear)$adj.r.squared,
         summary(black.int)$adj.r.squared)
)
kable(r_2)
```

* There was no preference between the two models, as the AICc values are within two units of each other.


### Temperature predictions and power analyses
```{r temperature predictions black carp}
## Build the models with three temperature metrics (simple model here)
black.annual <- lm(log(AAM)~AnnualTemp, data = black.clean)
black.cold <- lm(log(AAM)~ColdTemp, data = black.clean)
black.gdd <- lm(log(AAM)~average_gdd_0, data = black.clean)
black.water <- lm(log(AAM)~WaterTemp, data = black.clean)
black.waterC <- lm(log(AAM)~WaterCold, data = black.clean)


## Power analyses - annual
# calculate the coefficient of determination
coe.annual <- summary(black.annual)$adj.r.squared
pwr.f2.test(u = 1, v = 22 - 1 - 1, f2 = coe.annual/(1 -coe.annual),
            sig.level = 0.05)


## Power analyses - cold
# calculate the coefficient of determination
coe.cold <- summary(black.cold)$adj.r.squared
pwr.f2.test(u = 1, v = 22 - 1 - 1, f2 = coe.cold/(1 -coe.cold),
            sig.level = 0.05)


## Power analyses - gdd0
# calculate the coefficient of determination
coe.gdd <- summary(black.gdd)$adj.r.squared
pwr.f2.test(u = 1, v = 22 - 1 - 1, f2 = coe.gdd/(1 -coe.gdd),
            sig.level = 0.05)

pwr.f2.test(u = 1, f2 = coe.gdd/(1 -coe.gdd),
            sig.level = 0.05, power = 0.8)


## Power analyses - water annual
# calculate the coefficient of determination
coe.water <- summary(black.water)$adj.r.squared
pwr.f2.test(u = 1, v = 22 - 1 - 1, f2 = coe.water/(1 -coe.water),
            sig.level = 0.05)

pwr.f2.test(u = 1, f2 = coe.water/(1 -coe.water),
            sig.level = 0.05, power = 0.8)


## Power analyses - water cold
# calculate the coefficient of determination
coe.waterC <- summary(black.waterC)$adj.r.squared
pwr.f2.test(u = 1, v = 22 - 1 - 1, f2 = coe.waterC/(1 -coe.waterC),
            sig.level = 0.05)
```

* Power analyses suggested that our current sample size is sufficient enough to produce a strong statistical power for annual temperauture, cold temperature, gdd0 (?), and cold water temperature.

* Annual water?


### Black carp graphs with two conditions separated

We separated the black carp dataset into two based on conditions. Since there was no preference over the four models, we used the simple linear model on each set of the data.

```{r black carp graphs with two conditions}
## Annual temperature
ggplot(black.clean, aes(x = AnnualTemp, y = log(AAM), color = condition))+
  geom_point()+
  geom_smooth(method = "lm")+
  theme_bw()+
  labs(title = "Mean annual Temperature")


## Cold temperature
ggplot(black.clean, aes(x = ColdTemp, y = log(AAM), color = condition))+
  geom_point()+
  geom_smooth(method = "lm")+
  theme_bw()+
  labs(title = "Cold Quarter Temperature")


## GDD0
ggplot(black.clean, aes(x = average_gdd_0, y = log(AAM), color = condition))+
  geom_point()+
  geom_smooth(method = "lm")+
  theme_bw()+
  labs(title = "Growing degree day 0")


## Annual water temperature
ggplot(black.clean, aes(x = WaterTemp, y = log(AAM), color = condition))+
  geom_point()+
  geom_smooth(method = "lm")+
  theme_bw()+
  labs(title = "Growing degree day 0")


## Cold water temperature
ggplot(black.clean, aes(x = WaterCold, y = log(AAM), color = condition))+
  geom_point()+
  geom_smooth(method = "lm")+
  theme_bw()+
  labs(title = "Growing degree day 0")
```

Now that we have seen that the artificial condition data seems to have a larger spread, we would like to run a regression with only the natural conditions to see how much the R2 value can improve? Would it be similar to other Asian carp species (around 0.6)?

```{r black carp conditions}
## Separate into two data sets
black.natural <- black.clean[black.clean$condition == "natural",]
black.artificial <- black.clean[black.clean$condition == "artificial",]


## Run the models
black.annual.n <- lm(log(AAM)~AnnualTemp, data = black.natural)
black.cold.n <- lm(log(AAM)~ColdTemp, data = black.natural)
black.gdd.n <- lm(log(AAM)~average_gdd_0, data = black.natural)
black.water.n <- lm(log(AAM)~WaterTemp, data = black.natural)
black.waterC.n <- lm(log(AAM)~WaterCold, data = black.natural)


black.annual.a <- lm(log(AAM)~AnnualTemp, data = black.artificial)
black.cold.a <- lm(log(AAM)~ColdTemp, data = black.artificial)
black.gdd.a <- lm(log(AAM)~average_gdd_0, data = black.artificial)
black.water.a <- lm(log(AAM)~WaterTemp, data = black.artificial)
black.waterC.a <- lm(log(AAM)~WaterCold, data = black.artificial)


## Compare the model parameters
summary(black.annual.n)
summary(black.annual.a)

summary(black.cold.n)
summary(black.cold.a)

summary(black.gdd.n)
summary(black.gdd.a)

summary(black.water.n)
summary(black.water.a)

summary(black.waterC.n)
summary(black.waterC.a)


## Comparing R^2
r_2 <- data.frame(
  Temperature = c("Annual", "Cold", "GGD0", "Water", "Water cold"),
  Black.combined = c(0.31, 0.38, 0.27, 0.17, 0.29),
  Black.natural = c(summary(black.annual.n)$adj.r.squared,
                    summary(black.cold.n)$adj.r.squared,
                    summary(black.gdd.n)$adj.r.squared,
                    summary(black.water.n)$adj.r.squared,
                    summary(black.waterC.n)$adj.r.squared),
  Black.artificial = c(summary(black.annual.a)$adj.r.squared,
                       summary(black.cold.a)$adj.r.squared,
                       summary(black.gdd.a)$adj.r.squared,
                       summary(black.water.a)$adj.r.squared,
                       summary(black.waterC.a)$adj.r.squared),
  Asian.total = c(0.6, 0.57, 0.62, NA, NA)
)
kable(r_2)
```

* It turned out that removing the artificial conditions greatly increase the predictive power of the model for all five temperatures. While the artificial model alone did not even have a significant relationship.

* The resulting R2 (for the natural environments) is similar to that of the Asian carp.



## SECTION 2: Asian carp

1. We followed the same work flow for other Asian carp species. However, for Asian carp, we subsample at a distance of 250 km to avoid spatial autocorrelation.


### Data cleaning and matrices for results

```{r asian data}
## Look at the spatial codes for the current asian carp data
asian.carp.clean <- asian.carp %>% 
  filter(Condition %in% c("natural", "artificial"))

table(asian.carp.clean$Code_Str)
# Stratified sub-sampling gives 21 artificial and 17 natural conditions.

# Simple linear model - slope, intercept, p, r2, AICc, F stats
linear.results <- matrix(NA,1000,18)
colnames(linear.results) <- c("slope.a",
                                   "intercept.a",
                                   "p for slope.a",
                                   "r2.a",
                                   "AICc.a",
                                   "F stats.a",
                               "slope.c",
                                   "intercept.c",
                                   "p for slope.c",
                                   "r2.c",
                                   "AICc.c",
                                   "F stats.c",
                               "slope.g",
                                   "intercept.g",
                                   "p for slope.g",
                                   "r2.g",
                                   "AICc.g",
                                   "F stats.g")

# Interaction model - slope, intercept, p(slope), p(interaction), r2, AICc
int.results <- matrix(NA,1000,24)
colnames(int.results) <- c("slope.a",
                                   "intercept.a",
                                   "p for slope.a",
                                   "r2.a",
                                   "AICc.a",
                                   "F stats.a",
                                   "p additive term.a",
                                   "p interaction term.a",
                           "slope.c",
                                   "intercept.c",
                                   "p for slope.c",
                                   "r2.c",
                                   "AICc.c",
                                   "F stats.c",
                                   "p additive term.c",
                                   "p interaction term.c",
                           "slope.g",
                                   "intercept.g",
                                   "p for slope.g",
                                   "r2.g",
                                   "AICc.g",
                                   "F stats.g",
                                   "p additive term.g",
                                   "p interaction term.g")
```

* Stratified sub-sampling gives 21 artificial and 17 natural conditions.

### Define the models

```{r asian carp model}
# For 1000 iterations
for(i in 1:1000){
  sub <- asian.carp.clean %>% group_by(Code_Str) %>% sample_n(size=1)
  
  ## annual
  reg.linear.annual <- lm(log(AAM)~AnnualTemp, data = sub)
  reg.int.annual <- lm(log(AAM)~AnnualTemp*Condition, data = sub)
  
  # simple linear model
  linear.results[i,1]<-summary(reg.linear.annual)$coef[2,1] #slope
  linear.results[i,2]<-summary(reg.linear.annual)$coef[1,1] #intercept
  linear.results[i,3]<-summary(reg.linear.annual)$coef[2,4] #p-value
  linear.results[i,4]<-summary(reg.linear.annual)$adj.r.squared #r2
  linear.results[i,5]<-as.numeric(AICc(reg.linear.annual)) #AICc
  linear.results[i,6]<-summary(reg.linear.annual)$fstatistic[1] #F stats

  # interaction model
  int.results[i,1]<-summary(reg.int.annual)$coef[2,1] #slope
  int.results[i,2]<-summary(reg.int.annual)$coef[1,1] #intercept
  int.results[i,3]<-summary(reg.int.annual)$coef[2,4] #p(slope)
  int.results[i,4]<-summary(reg.int.annual)$adj.r.squared #r2
  int.results[i,5]<-as.numeric(AICc(reg.int.annual)) #AICc
  int.results[i,6]<-summary(reg.int.annual)$fstatistic[1] #F stats
  int.results[i,7]<-summary(reg.int.annual)$coef[3,4] #p(additive term)
  int.results[i,8]<-summary(reg.int.annual)$coef[4,4] #p(interaction term)


  ## cold
  reg.linear.cold <- lm(log(AAM)~ColdTemp, data = sub)
  reg.int.cold <- lm(log(AAM)~ColdTemp*Condition, data = sub)
  
  # simple linear model
  linear.results[i,7]<-summary(reg.linear.cold)$coef[2,1] #slope
  linear.results[i,8]<-summary(reg.linear.cold)$coef[1,1] #intercept
  linear.results[i,9]<-summary(reg.linear.cold)$coef[2,4] #p-value
  linear.results[i,10]<-summary(reg.linear.cold)$adj.r.squared #r2
  linear.results[i,11]<-as.numeric(AICc(reg.linear.cold)) #AICc
  linear.results[i,12]<-summary(reg.linear.cold)$fstatistic[1] #F stats

  # interaction model
  int.results[i,9]<-summary(reg.int.cold)$coef[2,1] #slope
  int.results[i,10]<-summary(reg.int.cold)$coef[1,1] #intercept
  int.results[i,11]<-summary(reg.int.cold)$coef[2,4] #p(slope)
  int.results[i,12]<-summary(reg.int.cold)$adj.r.squared #r2
  int.results[i,13]<-as.numeric(AICc(reg.int.cold)) #AICc
  int.results[i,14]<-summary(reg.int.cold)$fstatistic[1] #F stats
  int.results[i,15]<-summary(reg.int.cold)$coef[3,4] #p(additive term)
  int.results[i,16]<-summary(reg.int.cold)$coef[4,4] #p(interaction term)
  
  
  ## gdd0
  reg.linear.gdd <- lm(log(AAM)~average_gdd_0, data = sub)
  reg.int.gdd <- lm(log(AAM)~average_gdd_0*Condition, data = sub)
  
  # simple linear model
  linear.results[i,13]<-summary(reg.linear.gdd)$coef[2,1] #slope
  linear.results[i,14]<-summary(reg.linear.gdd)$coef[1,1] #intercept
  linear.results[i,15]<-summary(reg.linear.gdd)$coef[2,4] #p-value
  linear.results[i,16]<-summary(reg.linear.gdd)$adj.r.squared #r2
  linear.results[i,17]<-as.numeric(AICc(reg.linear.gdd)) #AICc
  linear.results[i,18]<-summary(reg.linear.gdd)$fstatistic[1] #F stats

  # interaction model
  int.results[i,17]<-summary(reg.int.gdd)$coef[2,1] #slope
  int.results[i,18]<-summary(reg.int.gdd)$coef[1,1] #intercept
  int.results[i,19]<-summary(reg.int.gdd)$coef[2,4] #p(slope)
  int.results[i,20]<-summary(reg.int.gdd)$adj.r.squared #r2
  int.results[i,21]<-as.numeric(AICc(reg.int.gdd)) #AICc
  int.results[i,22]<-summary(reg.int.gdd)$fstatistic[1] #F stats
  int.results[i,23]<-summary(reg.int.gdd)$coef[3,4] #p(additive term)
  int.results[i,24]<-summary(reg.int.cold)$coef[4,4] #p(interaction term)
}
```

### Check the slopes and additive/interaction terms

Now we need to:

1. Check if the additive or interaction term is significant.

2. Check if the relationship is significant between age at maturity and temperature (significant slope).

```{r asian slope and intercept}
## Additive/interaction term
# annual
mean(unique(int.results[,"p additive term.a"]))
table(int.results[,"p additive term.a"] < 0.05)
mean(unique(int.results[,"p interaction term.a"]))
table(int.results[,"p interaction term.a"] < 0.05)

# cold
mean(unique(int.results[,"p additive term.c"]))
table(int.results[,"p additive term.c"] < 0.05)
mean(unique(int.results[,"p interaction term.c"]))
table(int.results[,"p interaction term.c"] < 0.05)

# gdd
mean(unique(int.results[,"p additive term.g"]))
table(int.results[,"p additive term.g"] < 0.05)
mean(unique(int.results[,"p interaction term.g"]))
table(int.results[,"p interaction term.g"] < 0.05)

# We found no siginficant additive or interaction term


### Now let's look at the simple regression model:

## p-value for the slope (also is the p-value for the entire model)
# annual
mean(unique(linear.results[,"p for slope.a"]))
table(linear.results[,"p for slope.a"] < 0.05)

# cold
mean(unique(linear.results[,"p for slope.c"]))
table(linear.results[,"p for slope.c"] < 0.05)

# gdd
mean(unique(linear.results[,"p for slope.g"]))
table(linear.results[,"p for slope.g"] < 0.05)


## F statistics
mean(unique(linear.results[,"F stats.a"]))
mean(unique(linear.results[,"F stats.c"]))
mean(unique(linear.results[,"F stats.g"]))


## Slope
mean(unique(linear.results[,"slope.a"]))
mean(unique(linear.results[,"slope.c"]))
mean(unique(linear.results[,"slope.g"]))


## Intercept
mean(unique(linear.results[,"intercept.a"]))
mean(unique(linear.results[,"intercept.c"]))
mean(unique(linear.results[,"intercept.g"]))
```
* No significant additive or interaction term in the multiple linear regression model. As a result, we should use the simple regression model.

* In a simple regression model, significant slope using all three temperatures.


### Compare the R2

```{r asian r2}
# annual
r2annual <- data.frame(
  Model = c("Simple linear", "Interaction"),
  R2 = c(mean(unique(linear.results[,4])),
         mean(unique(int.results[,4])))
)
kable(r2annual)

# cold
r2cold <- data.frame(
  Model = c("Simple linear", "Interaction"),
  R2 = c(mean(unique(linear.results[,10])),
         mean(unique(int.results[,12])))
)
kable(r2cold)

# gdd
r2gdd <- data.frame(
  Model = c("Simple linear", "Interaction"),
  R2 = c(mean(unique(linear.results[,16])),
         mean(unique(int.results[,20])))
)
kable(r2gdd)
```


### Compare among the temperatures (SIMPLE MODEL ONLY)

```{r asian temperature compare}
## Compare the R2 for three temperatures
r2 <- data.frame(
  Temperature = c("Annual", "Cold", "GDD0"),
  R2 = c(mean(unique(linear.results[,"r2.a"])),
         mean(unique(linear.results[,"r2.c"])),
         mean(unique(linear.results[,"r2.g"])))
)
kable(r2)


## Check the AICc scores and akaike weights in ONLY LINEAR MODEL
weight.matrix <- matrix(NA, 1000, 3)
count <- numeric(0)

for (i in 1:1000) {
  # Create a list of the aic values of the current iteration
  aic_value <- c(linear.results[i,5], linear.results[i,11],
                 linear.results[i,17])
  
  ## check the akaike weights
  weight <- compute_akaike_weights(aic_value)
  weight.matrix[i,c(1,2,3)] <- round(weight[c(1,2,3)],3)
  
  ## check the AICc scores
  indexing <- compare_aic_scores(aic_value)
  if (indexing != -999) {
    count <- c(count, indexing)
  }
}

summary(weight.matrix)
table(count)
```

* With the simple linear model, around 30% of the time when using GDD is preferred over using ColdTemp.


### Two conditions separated

```{r asian separate conditions}
## Separate the two conditions
asian.natural <- asian.carp.clean[asian.carp.clean$Condition == "natural",]
asian.artificial <- asian.carp.clean[asian.carp.clean$Condition == "artificial",]

# Check sp
table(asian.natural$Code_Str)
table(asian.artificial$Code_Str)

# store the results
r2.natural <- matrix(NA, 1000, 3)
r2.artificial <- matrix(NA, 1000, 3)

p.natural <- matrix(NA, 1000, 3)
p.artificial <- matrix(NA, 1000, 3)

F.natural <- matrix(NA, 1000, 3)
F.artificial <- matrix(NA, 1000, 3)


## Separating environment for 1000 iterations
for(i in 1:1000){
  
  ## natural
  sub.n <- asian.natural %>% group_by(Code_Str) %>% sample_n(size=1)
  
  # models
  reg.annual <- lm(log(AAM)~AnnualTemp, data = sub.n)
  reg.cold <- lm(log(AAM)~ColdTemp, data = sub.n)
  reg.gdd <- lm(log(AAM)~average_gdd_0, data = sub.n)

  # r2
  r2.natural[i,1]<-summary(reg.annual)$adj.r.squared
  r2.natural[i,2]<-summary(reg.cold)$adj.r.squared
  r2.natural[i,3]<-summary(reg.gdd)$adj.r.squared
  
  # p
  p.natural[i,1]<-summary(reg.annual)$coef[2,4]
  p.natural[i,2]<-summary(reg.cold)$coef[2,4]
  p.natural[i,3]<-summary(reg.gdd)$coef[2,4]
  
  # F
  F.natural[i,1] <- summary(reg.annual)$fstatistic[1]
  F.natural[i,2] <- summary(reg.cold)$fstatistic[1]
  F.natural[i,3] <- summary(reg.gdd)$fstatistic[1]
  
  
  ## artificial
  sub.a <- asian.artificial %>% group_by(Code_Str) %>% sample_n(size=1)

  # models
  reg.annual <- lm(log(AAM)~AnnualTemp, data = sub.a)
  reg.cold <- lm(log(AAM)~ColdTemp, data = sub.a)
  reg.gdd <- lm(log(AAM)~average_gdd_0, data = sub.a)

  # r2
  r2.artificial[i,1]<-summary(reg.annual)$adj.r.squared
  r2.artificial[i,2]<-summary(reg.cold)$adj.r.squared
  r2.artificial[i,3]<-summary(reg.gdd)$adj.r.squared
  
  # p
  p.artificial[i,1]<-summary(reg.annual)$coef[2,4]
  p.artificial[i,2]<-summary(reg.cold)$coef[2,4]
  p.artificial[i,3]<-summary(reg.gdd)$coef[2,4]
  
  # F
  F.artificial[i,1] <- summary(reg.annual)$fstatistic[1]
  F.artificial[i,2] <- summary(reg.cold)$fstatistic[1]
  F.artificial[i,3] <- summary(reg.gdd)$fstatistic[1]
}


## Compare the R2 for three temperatures
r2 <- data.frame(
  Temperature = c("Annual", "Cold", "GDD0"),
  Natural.R2 = c(mean(unique(r2.natural[,1])),
                 mean(unique(r2.natural[,2])),
                 mean(unique(r2.natural[,3]))),
  Artificial.R2 = c(mean(unique(r2.artificial[,1])),
                    mean(unique(r2.artificial[,2])),
                    mean(unique(r2.artificial[,3])))
)
kable(r2)


## Compare the F stats
Fstats <- data.frame(
  Temperature = c("Annual", "Cold", "GDD0"),
  Natural.F = c(mean(unique(F.natural[,1])),
                 mean(unique(F.natural[,2])),
                 mean(unique(F.natural[,3]))),
  Artificial.F = c(mean(unique(F.artificial[,1])),
                    mean(unique(F.artificial[,2])),
                    mean(unique(F.artificial[,3])))
)
kable(Fstats)


## Check the p values
table(p.natural[,1] < 0.05) # natural, annual
table(p.natural[,2] < 0.05) # natural, cold
table(p.natural[,3] < 0.05) # natural, gdd
table(p.artificial[,1] < 0.05) # artificial, annual
table(p.artificial[,2] < 0.05) # artificial, cold
table(p.artificial[,3] < 0.05) # artificial, gdd
```

* Natural and Artificial have similar R2 values when separated.



## SECTION 3: Grass carp

### Data cleaning and matrices for results

```{r grass data}
Grass.clean <- Grass %>% 
  filter(Condition %in% c("natural", "artificial"))

table(Grass.clean$Code_Str)


# Simple linear model - slope, intercept, p, r2, AICc, F stats
linear.results <- matrix(NA,1000,18)
colnames(linear.results) <- c("slope.a",
                                   "intercept.a",
                                   "p for slope.a",
                                   "r2.a",
                                   "AICc.a",
                                   "F stats.a",
                               "slope.c",
                                   "intercept.c",
                                   "p for slope.c",
                                   "r2.c",
                                   "AICc.c",
                                   "F stats.c",
                               "slope.g",
                                   "intercept.g",
                                   "p for slope.g",
                                   "r2.g",
                                   "AICc.g",
                                   "F stats.g")

# Interaction model - slope, intercept, p(slope), p(interaction), r2, AICc
int.results <- matrix(NA,1000,24)
colnames(int.results) <- c("slope.a",
                                   "intercept.a",
                                   "p for slope.a",
                                   "r2.a",
                                   "AICc.a",
                                   "F stats.a",
                                   "p additive term.a",
                                   "p interaction term.a",
                           "slope.c",
                                   "intercept.c",
                                   "p for slope.c",
                                   "r2.c",
                                   "AICc.c",
                                   "F stats.c",
                                   "p additive term.c",
                                   "p interaction term.c",
                           "slope.g",
                                   "intercept.g",
                                   "p for slope.g",
                                   "r2.g",
                                   "AICc.g",
                                   "F stats.g",
                                   "p additive term.g",
                                   "p interaction term.g")
```


### Define the models

```{r grass carp model}
# For 1000 iterations
for(i in 1:1000){
  sub <- Grass.clean %>% group_by(Code_Str) %>% sample_n(size=1)
  
  ## annual
  reg.linear.annual <- lm(log(AAM)~AnnualTemp, data = sub)
  reg.int.annual <- lm(log(AAM)~AnnualTemp*Condition, data = sub)
  
  # simple linear model
  linear.results[i,1]<-summary(reg.linear.annual)$coef[2,1] #slope
  linear.results[i,2]<-summary(reg.linear.annual)$coef[1,1] #intercept
  linear.results[i,3]<-summary(reg.linear.annual)$coef[2,4] #p-value
  linear.results[i,4]<-summary(reg.linear.annual)$adj.r.squared #r2
  linear.results[i,5]<-as.numeric(AICc(reg.linear.annual)) #AICc
  linear.results[i,6]<-summary(reg.linear.annual)$fstatistic[1] #F stats

  # interaction model
  int.results[i,1]<-summary(reg.int.annual)$coef[2,1] #slope
  int.results[i,2]<-summary(reg.int.annual)$coef[1,1] #intercept
  int.results[i,3]<-summary(reg.int.annual)$coef[2,4] #p(slope)
  int.results[i,4]<-summary(reg.int.annual)$adj.r.squared #r2
  int.results[i,5]<-as.numeric(AICc(reg.int.annual)) #AICc
  int.results[i,6]<-summary(reg.int.annual)$fstatistic[1] #F stats
  int.results[i,7]<-summary(reg.int.annual)$coef[3,4] #p(additive term)
  int.results[i,8]<-summary(reg.int.annual)$coef[4,4] #p(interaction term)


  ## cold
  reg.linear.cold <- lm(log(AAM)~ColdTemp, data = sub)
  reg.int.cold <- lm(log(AAM)~ColdTemp*Condition, data = sub)
  
  # simple linear model
  linear.results[i,7]<-summary(reg.linear.cold)$coef[2,1] #slope
  linear.results[i,8]<-summary(reg.linear.cold)$coef[1,1] #intercept
  linear.results[i,9]<-summary(reg.linear.cold)$coef[2,4] #p-value
  linear.results[i,10]<-summary(reg.linear.cold)$adj.r.squared #r2
  linear.results[i,11]<-as.numeric(AICc(reg.linear.cold)) #AICc
  linear.results[i,12]<-summary(reg.linear.cold)$fstatistic[1] #F stats

  # interaction model
  int.results[i,9]<-summary(reg.int.cold)$coef[2,1] #slope
  int.results[i,10]<-summary(reg.int.cold)$coef[1,1] #intercept
  int.results[i,11]<-summary(reg.int.cold)$coef[2,4] #p(slope)
  int.results[i,12]<-summary(reg.int.cold)$adj.r.squared #r2
  int.results[i,13]<-as.numeric(AICc(reg.int.cold)) #AICc
  int.results[i,14]<-summary(reg.int.cold)$fstatistic[1] #F stats
  int.results[i,15]<-summary(reg.int.cold)$coef[3,4] #p(additive term)
  int.results[i,16]<-summary(reg.int.cold)$coef[4,4] #p(interaction term)
  
  
  ## gdd0
  reg.linear.gdd <- lm(log(AAM)~average_gdd_0, data = sub)
  reg.int.gdd <- lm(log(AAM)~average_gdd_0*Condition, data = sub)
  
  # simple linear model
  linear.results[i,13]<-summary(reg.linear.gdd)$coef[2,1] #slope
  linear.results[i,14]<-summary(reg.linear.gdd)$coef[1,1] #intercept
  linear.results[i,15]<-summary(reg.linear.gdd)$coef[2,4] #p-value
  linear.results[i,16]<-summary(reg.linear.gdd)$adj.r.squared #r2
  linear.results[i,17]<-as.numeric(AICc(reg.linear.gdd)) #AICc
  linear.results[i,18]<-summary(reg.linear.gdd)$fstatistic[1] #F stats

  # interaction model
  int.results[i,17]<-summary(reg.int.gdd)$coef[2,1] #slope
  int.results[i,18]<-summary(reg.int.gdd)$coef[1,1] #intercept
  int.results[i,19]<-summary(reg.int.gdd)$coef[2,4] #p(slope)
  int.results[i,20]<-summary(reg.int.gdd)$adj.r.squared #r2
  int.results[i,21]<-as.numeric(AICc(reg.int.gdd)) #AICc
  int.results[i,22]<-summary(reg.int.gdd)$fstatistic[1] #F stats
  int.results[i,23]<-summary(reg.int.gdd)$coef[3,4] #p(additive term)
  int.results[i,24]<-summary(reg.int.cold)$coef[4,4] #p(interaction term)
}
```


### Check the slopes and additive/interaction terms

Now we need to:

1. Check if the additive or interaction term is significant.

2. Check if the relationship is significant between age at maturity and temperature (significant slope).

```{r grass slope and intercept}
## Additive/interaction term
# annual
mean(unique(int.results[,"p additive term.a"]))
table(int.results[,"p additive term.a"] < 0.05)
mean(unique(int.results[,"p interaction term.a"]))
table(int.results[,"p interaction term.a"] < 0.05)

# cold
mean(unique(int.results[,"p additive term.c"]))
table(int.results[,"p additive term.c"] < 0.05)
mean(unique(int.results[,"p interaction term.c"]))
table(int.results[,"p interaction term.c"] < 0.05)

# gdd
mean(unique(int.results[,"p additive term.g"]))
table(int.results[,"p additive term.g"] < 0.05)
mean(unique(int.results[,"p interaction term.g"]))
table(int.results[,"p interaction term.g"] < 0.05)


### Now let's look at the simple regression model:

## p-value for the slope (also is the p-value for the entire model)
# annual
mean(unique(linear.results[,"p for slope.a"]))
table(linear.results[,"p for slope.a"] < 0.05)

# cold
mean(unique(linear.results[,"p for slope.c"]))
table(linear.results[,"p for slope.c"] < 0.05)

# gdd
mean(unique(linear.results[,"p for slope.g"]))
table(linear.results[,"p for slope.g"] < 0.05)

## F statistics
mean(unique(linear.results[,"F stats.a"]))
mean(unique(linear.results[,"F stats.c"]))
mean(unique(linear.results[,"F stats.g"]))
```

* No significant additive or interaction term in the multiple linear model.

* Significant slope.


### Compare the R2

```{r grass r2}
# annual
r2annual <- data.frame(
  Model = c("Simple linear", "Interaction"),
  R2 = c(mean(unique(linear.results[,5])),
         mean(unique(int.results[,5])))
)
kable(r2annual)

# cold
r2cold <- data.frame(
  Model = c("Simple linear", "Interaction"),
  R2 = c(mean(unique(linear.results[,11])),
         mean(unique(int.results[,11])))
)
kable(r2cold)

# gdd
r2gdd <- data.frame(
  Model = c("Simple linear", "Interaction"),
  R2 = c(mean(unique(linear.results[,17])),
         mean(unique(int.results[,17])))
)
kable(r2gdd)
```


### Compare among three temperatures (SIMPLE MODEL ONLY)

```{r grass temperature compare}
## Compare the R2 for three temperatures
r2 <- data.frame(
  Temperature = c("Annual", "Cold", "GDD0"),
  R2 = c(mean(unique(linear.results[,"r2.a"])),
         mean(unique(linear.results[,"r2.c"])),
         mean(unique(linear.results[,"r2.g"])))
)
kable(r2)


## Check the AICc scores and akaike weights in ONLY LINEAR MODEL
weight.matrix <- matrix(NA, 1000, 3)
count <- numeric(0)

for (i in 1:1000) {
  # Create a list of the aic values of the current iteration
  aic_value <- c(linear.results[i,5], linear.results[i,11],
                 linear.results[i,17])
  
  ## check the akaike weights
  weight <- compute_akaike_weights(aic_value)
  weight.matrix[i,c(1,2,3)] <- round(weight[c(1,2,3)],3)
  
  ## check the AICc scores
  indexing <- compare_aic_scores(aic_value)
  if (indexing != -999) {
    count <- c(count, indexing)
  }
}

summary(weight.matrix)
table(count)
```

* Cold temperature did not show any preference over Annual temperature. No significant differences.


### Two conditions separated
```{r grass separate conditions}
## Separate the two conditions
grass.natural <- Grass.clean[Grass.clean$Condition == "natural",]
grass.artificial <- Grass.clean[Grass.clean$Condition == "artificial",]

# Check sp
table(grass.natural$Code_Str)
table(grass.artificial$Code_Str)

# store the results
r2.natural <- matrix(NA, 1000, 3)
r2.artificial <- matrix(NA, 1000, 3)

p.natural <- matrix(NA, 1000, 3)
p.artificial <- matrix(NA, 1000, 3)

F.natural <- matrix(NA, 1000, 3)
F.artificial <- matrix(NA, 1000, 3)


## Separating environment for 1000 iterations
for(i in 1:1000){
  
  ## natural
  sub.n <- grass.natural %>% group_by(Code_Str) %>% sample_n(size=1)
  
  # models
  reg.annual <- lm(log(AAM)~AnnualTemp, data = sub.n)
  reg.cold <- lm(log(AAM)~ColdTemp, data = sub.n)
  reg.gdd <- lm(log(AAM)~average_gdd_0, data = sub.n)

  # r2
  r2.natural[i,1]<-summary(reg.annual)$adj.r.squared
  r2.natural[i,2]<-summary(reg.cold)$adj.r.squared
  r2.natural[i,3]<-summary(reg.gdd)$adj.r.squared
  
  # p
  p.natural[i,1]<-summary(reg.annual)$coef[2,4]
  p.natural[i,2]<-summary(reg.cold)$coef[2,4]
  p.natural[i,3]<-summary(reg.gdd)$coef[2,4]
  
  # F
  F.natural[i,1] <- summary(reg.annual)$fstatistic[1]
  F.natural[i,2] <- summary(reg.cold)$fstatistic[1]
  F.natural[i,3] <- summary(reg.gdd)$fstatistic[1]
  
  
  ## artificial
  sub.a <- grass.artificial %>% group_by(Code_Str) %>% sample_n(size=1)

  # models
  reg.annual <- lm(log(AAM)~AnnualTemp, data = sub.a)
  reg.cold <- lm(log(AAM)~ColdTemp, data = sub.a)
  reg.gdd <- lm(log(AAM)~average_gdd_0, data = sub.a)

  # r2
  r2.artificial[i,1]<-summary(reg.annual)$adj.r.squared
  r2.artificial[i,2]<-summary(reg.cold)$adj.r.squared
  r2.artificial[i,3]<-summary(reg.gdd)$adj.r.squared
  
  # p
  p.artificial[i,1]<-summary(reg.annual)$coef[2,4]
  p.artificial[i,2]<-summary(reg.cold)$coef[2,4]
  p.artificial[i,3]<-summary(reg.gdd)$coef[2,4]
  
  # F
  F.artificial[i,1] <- summary(reg.annual)$fstatistic[1]
  F.artificial[i,2] <- summary(reg.cold)$fstatistic[1]
  F.artificial[i,3] <- summary(reg.gdd)$fstatistic[1]
}


## Compare the R2 for three temperatures
r2 <- data.frame(
  Temperature = c("Annual", "Cold", "GDD0"),
  Natural.R2 = c(mean(unique(r2.natural[,1])),
                 mean(unique(r2.natural[,2])),
                 mean(unique(r2.natural[,3]))),
  Artificial.R2 = c(mean(unique(r2.artificial[,1])),
                    mean(unique(r2.artificial[,2])),
                    mean(unique(r2.artificial[,3])))
)
kable(r2)

sub.n <- grass.artificial %>% group_by(Code_Str) %>% sample_n(size=1)
reg.annual <- lm(log(AAM)~AnnualTemp, data = sub.n)
summary(reg.annual)


## Compare the F stats
Fstats <- data.frame(
  Temperature = c("Annual", "Cold", "GDD0"),
  Natural.F = c(mean(unique(F.natural[,1])),
                 mean(unique(F.natural[,2])),
                 mean(unique(F.natural[,3]))),
  Artificial.F = c(mean(unique(F.artificial[,1])),
                    mean(unique(F.artificial[,2])),
                    mean(unique(F.artificial[,3])))
)
kable(Fstats)


## Check the p values
table(p.natural[,1] < 0.05) # natural, annual
table(p.natural[,2] < 0.05) # natural, cold
table(p.natural[,3] < 0.05) # natural, gdd
table(p.artificial[,1] < 0.05) # artificial, annual
table(p.artificial[,2] < 0.05) # artificial, cold
table(p.artificial[,3] < 0.05) # artificial, gdd
```



## SECTION 4: Bighead and silver carp

### Data cleaning and matrices for results

```{r bs data}
Big.sil.clean <- Big.sil %>% 
  filter(Condition %in% c("natural", "artificial"))

table(Big.sil.clean$Code)


# Simple linear model - slope, intercept, p, r2, AICc, F stats
linear.results <- matrix(NA,1000,18)
colnames(linear.results) <- c("slope.a",
                                   "intercept.a",
                                   "p for slope.a",
                                   "r2.a",
                                   "AICc.a",
                                   "F stats.a",
                               "slope.c",
                                   "intercept.c",
                                   "p for slope.c",
                                   "r2.c",
                                   "AICc.c",
                                   "F stats.c",
                               "slope.g",
                                   "intercept.g",
                                   "p for slope.g",
                                   "r2.g",
                                   "AICc.g",
                                   "F stats.g")

# Interaction model - slope, intercept, p(slope), p(interaction), r2, AICc
int.results <- matrix(NA,1000,24)
colnames(int.results) <- c("slope.a",
                                   "intercept.a",
                                   "p for slope.a",
                                   "r2.a",
                                   "AICc.a",
                                   "F stats.a",
                                   "p additive term.a",
                                   "p interaction term.a",
                           "slope.c",
                                   "intercept.c",
                                   "p for slope.c",
                                   "r2.c",
                                   "AICc.c",
                                   "F stats.c",
                                   "p additive term.c",
                                   "p interaction term.c",
                           "slope.g",
                                   "intercept.g",
                                   "p for slope.g",
                                   "r2.g",
                                   "AICc.g",
                                   "F stats.g",
                                   "p additive term.g",
                                   "p interaction term.g")
```
* Stratified sub-sampling gives 13 artificial points and 10 natural points.


### Define the models

```{r bs carp model}
# For 1000 iterations
for(i in 1:1000){
  sub <- Big.sil.clean %>% group_by(Code_Str) %>% sample_n(size=1)
  
  ## annual
  reg.linear.annual <- lm(log(AAM)~AnnualTemp, data = sub)
  reg.int.annual <- lm(log(AAM)~AnnualTemp*Condition, data = sub)
  
  # simple linear model
  linear.results[i,1]<-summary(reg.linear.annual)$coef[2,1] #slope
  linear.results[i,2]<-summary(reg.linear.annual)$coef[1,1] #intercept
  linear.results[i,3]<-summary(reg.linear.annual)$coef[2,4] #p-value
  linear.results[i,4]<-summary(reg.linear.annual)$adj.r.squared #r2
  linear.results[i,5]<-as.numeric(AICc(reg.linear.annual)) #AICc
  linear.results[i,6]<-summary(reg.linear.annual)$fstatistic[1] #F stats

  # interaction model
  int.results[i,1]<-summary(reg.int.annual)$coef[2,1] #slope
  int.results[i,2]<-summary(reg.int.annual)$coef[1,1] #intercept
  int.results[i,3]<-summary(reg.int.annual)$coef[2,4] #p(slope)
  int.results[i,4]<-summary(reg.int.annual)$adj.r.squared #r2
  int.results[i,5]<-as.numeric(AICc(reg.int.annual)) #AICc
  int.results[i,6]<-summary(reg.int.annual)$fstatistic[1] #F stats
  int.results[i,7]<-summary(reg.int.annual)$coef[3,4] #p(additive term)
  int.results[i,8]<-summary(reg.int.annual)$coef[4,4] #p(interaction term)


  ## cold
  reg.linear.cold <- lm(log(AAM)~ColdTemp, data = sub)
  reg.int.cold <- lm(log(AAM)~ColdTemp*Condition, data = sub)
  
  # simple linear model
  linear.results[i,7]<-summary(reg.linear.cold)$coef[2,1] #slope
  linear.results[i,8]<-summary(reg.linear.cold)$coef[1,1] #intercept
  linear.results[i,9]<-summary(reg.linear.cold)$coef[2,4] #p-value
  linear.results[i,10]<-summary(reg.linear.cold)$adj.r.squared #r2
  linear.results[i,11]<-as.numeric(AICc(reg.linear.cold)) #AICc
  linear.results[i,12]<-summary(reg.linear.cold)$fstatistic[1] #F stats

  # interaction model
  int.results[i,9]<-summary(reg.int.cold)$coef[2,1] #slope
  int.results[i,10]<-summary(reg.int.cold)$coef[1,1] #intercept
  int.results[i,11]<-summary(reg.int.cold)$coef[2,4] #p(slope)
  int.results[i,12]<-summary(reg.int.cold)$adj.r.squared #r2
  int.results[i,13]<-as.numeric(AICc(reg.int.cold)) #AICc
  int.results[i,14]<-summary(reg.int.cold)$fstatistic[1] #F stats
  int.results[i,15]<-summary(reg.int.cold)$coef[3,4] #p(additive term)
  int.results[i,16]<-summary(reg.int.cold)$coef[4,4] #p(interaction term)
  
  
  ## gdd0
  reg.linear.gdd <- lm(log(AAM)~average_gdd_0, data = sub)
  reg.int.gdd <- lm(log(AAM)~average_gdd_0*Condition, data = sub)
  
  # simple linear model
  linear.results[i,13]<-summary(reg.linear.gdd)$coef[2,1] #slope
  linear.results[i,14]<-summary(reg.linear.gdd)$coef[1,1] #intercept
  linear.results[i,15]<-summary(reg.linear.gdd)$coef[2,4] #p-value
  linear.results[i,16]<-summary(reg.linear.gdd)$adj.r.squared #r2
  linear.results[i,17]<-as.numeric(AICc(reg.linear.gdd)) #AICc
  linear.results[i,18]<-summary(reg.linear.gdd)$fstatistic[1] #F stats

  # interaction model
  int.results[i,17]<-summary(reg.int.gdd)$coef[2,1] #slope
  int.results[i,18]<-summary(reg.int.gdd)$coef[1,1] #intercept
  int.results[i,19]<-summary(reg.int.gdd)$coef[2,4] #p(slope)
  int.results[i,20]<-summary(reg.int.gdd)$adj.r.squared #r2
  int.results[i,21]<-as.numeric(AICc(reg.int.gdd)) #AICc
  int.results[i,22]<-summary(reg.int.gdd)$fstatistic[1] #F stats
  int.results[i,23]<-summary(reg.int.gdd)$coef[3,4] #p(additive term)
  int.results[i,24]<-summary(reg.int.cold)$coef[4,4] #p(interaction term)
}
```


### Check the slopes and additive/interaction terms

Now we need to:

1. Check if the additive or interaction term is significant.

2. Check if the relationship is significant between age at maturity and temperature (significant slope).

```{r bs slope and intercept}
## Additive/interaction term
# annual
mean(unique(int.results[,"p additive term.a"]))
table(int.results[,"p additive term.a"] < 0.05)
mean(unique(int.results[,"p interaction term.a"]))
table(int.results[,"p interaction term.a"] < 0.05)

# cold
mean(unique(int.results[,"p additive term.c"]))
table(int.results[,"p additive term.c"] < 0.05)
mean(unique(int.results[,"p interaction term.c"]))
table(int.results[,"p interaction term.c"] < 0.05)

# gdd
mean(unique(int.results[,"p additive term.g"]))
table(int.results[,"p additive term.g"] < 0.05)
mean(unique(int.results[,"p interaction term.g"]))
table(int.results[,"p interaction term.g"] < 0.05)


### Now let's look at the simple regression model:

## p-value for the slope (also is the p-value for the entire model)
# annual
mean(unique(linear.results[,"p for slope.a"]))
table(linear.results[,"p for slope.a"] < 0.05)

# cold
mean(unique(linear.results[,"p for slope.c"]))
table(linear.results[,"p for slope.c"] < 0.05)

# gdd
mean(unique(linear.results[,"p for slope.g"]))
table(linear.results[,"p for slope.g"] < 0.05)

## F statistics
mean(unique(linear.results[,"F stats.a"]))
mean(unique(linear.results[,"F stats.c"]))
mean(unique(linear.results[,"F stats.g"]))
```

* No significant additive or interaction term.

* Significant slope.


### Compare the R2

```{r bs r2}
# annual
r2annual <- data.frame(
  Model = c("Simple linear", "Interaction"),
  R2 = c(mean(unique(linear.results[,5])),
         mean(unique(int.results[,5])))
)
kable(r2annual)

# cold
r2cold <- data.frame(
  Model = c("Simple linear", "Interaction"),
  R2 = c(mean(unique(linear.results[,11])),
         mean(unique(int.results[,11])))
)
kable(r2cold)

# gdd
r2gdd <- data.frame(
  Model = c("Simple linear", "Interaction"),
  R2 = c(mean(unique(linear.results[,17])),
         mean(unique(int.results[,17])))
)
kable(r2gdd)
```


### Compare between annual and cold (SIMPLE MODEL ONLY)

```{r bs temperature compare}
## Compare the R2 for three temperatures
r2 <- data.frame(
  Temperature = c("Annual", "Cold", "GDD0"),
  R2 = c(mean(unique(linear.results[,"r2.a"])),
         mean(unique(linear.results[,"r2.c"])),
         mean(unique(linear.results[,"r2.g"])))
)
kable(r2)


## Check the AICc scores and akaike weights in ONLY LINEAR MODEL
weight.matrix <- matrix(NA, 1000, 3)
count <- numeric(0)

for (i in 1:1000) {
  # Create a list of the aic values of the current iteration
  aic_value <- c(linear.results[i,5], linear.results[i,11],
                 linear.results[i,17])
  
  ## check the akaike weights
  weight <- compute_akaike_weights(aic_value)
  weight.matrix[i,c(1,2,3)] <- round(weight[c(1,2,3)],3)
  
  ## check the AICc scores
  indexing <- compare_aic_scores(aic_value)
  if (indexing != -999) {
    count <- c(count, indexing)
  }
}

summary(weight.matrix)
table(count)
```

* For bs, using annual temperature is always better than using the cold temperature (difference in AIC > 2) for all models. This was also explained by lower R^2 values for the cold temperature models.


### Two conditions separated
```{r bs separate conditions}
## Separate the two conditions
bs.natural <- Big.sil.clean[Big.sil.clean$Condition == "natural",]
bs.artificial <- Big.sil.clean[Big.sil.clean$Condition == "artificial",]

# Check sp
table(bs.natural$Code_Str)
table(bs.artificial$Code_Str)

# store the results
r2.natural <- matrix(NA, 1000, 3)
r2.artificial <- matrix(NA, 1000, 3)

p.natural <- matrix(NA, 1000, 3)
p.artificial <- matrix(NA, 1000, 3)

F.natural <- matrix(NA, 1000, 3)
F.artificial <- matrix(NA, 1000, 3)


## Separating environment for 1000 iterations
for(i in 1:1000){
  
  ## natural
  sub.n <- bs.natural %>% group_by(Code_Str) %>% sample_n(size=1)
  
  # models
  reg.annual <- lm(log(AAM)~AnnualTemp, data = sub.n)
  reg.cold <- lm(log(AAM)~ColdTemp, data = sub.n)
  reg.gdd <- lm(log(AAM)~average_gdd_0, data = sub.n)

  # r2
  r2.natural[i,1]<-summary(reg.annual)$adj.r.squared
  r2.natural[i,2]<-summary(reg.cold)$adj.r.squared
  r2.natural[i,3]<-summary(reg.gdd)$adj.r.squared
  
  # p
  p.natural[i,1]<-summary(reg.annual)$coef[2,4]
  p.natural[i,2]<-summary(reg.cold)$coef[2,4]
  p.natural[i,3]<-summary(reg.gdd)$coef[2,4]
  
  # F
  F.natural[i,1] <- summary(reg.annual)$fstatistic[1]
  F.natural[i,2] <- summary(reg.cold)$fstatistic[1]
  F.natural[i,3] <- summary(reg.gdd)$fstatistic[1]
  
  
  ## artificial
  sub.a <- bs.artificial %>% group_by(Code_Str) %>% sample_n(size=1)

  # models
  reg.annual <- lm(log(AAM)~AnnualTemp, data = sub.a)
  reg.cold <- lm(log(AAM)~ColdTemp, data = sub.a)
  reg.gdd <- lm(log(AAM)~average_gdd_0, data = sub.a)

  # r2
  r2.artificial[i,1]<-summary(reg.annual)$adj.r.squared
  r2.artificial[i,2]<-summary(reg.cold)$adj.r.squared
  r2.artificial[i,3]<-summary(reg.gdd)$adj.r.squared
  
  # p
  p.artificial[i,1]<-summary(reg.annual)$coef[2,4]
  p.artificial[i,2]<-summary(reg.cold)$coef[2,4]
  p.artificial[i,3]<-summary(reg.gdd)$coef[2,4]
  
  # F
  F.artificial[i,1] <- summary(reg.annual)$fstatistic[1]
  F.artificial[i,2] <- summary(reg.cold)$fstatistic[1]
  F.artificial[i,3] <- summary(reg.gdd)$fstatistic[1]
}


## Compare the R2 for three temperatures
r2 <- data.frame(
  Temperature = c("Annual", "Cold", "GDD0"),
  Natural.R2 = c(mean(unique(r2.natural[,1])),
                 mean(unique(r2.natural[,2])),
                 mean(unique(r2.natural[,3]))),
  Artificial.R2 = c(mean(unique(r2.artificial[,1])),
                    mean(unique(r2.artificial[,2])),
                    mean(unique(r2.artificial[,3])))
)
kable(r2)


## Compare the F stats
Fstats <- data.frame(
  Temperature = c("Annual", "Cold", "GDD0"),
  Natural.F = c(mean(unique(F.natural[,1])),
                 mean(unique(F.natural[,2])),
                 mean(unique(F.natural[,3]))),
  Artificial.F = c(mean(unique(F.artificial[,1])),
                    mean(unique(F.artificial[,2])),
                    mean(unique(F.artificial[,3])))
)
kable(Fstats)


## Check the p values
table(p.natural[,1] < 0.05) # natural, annual
table(p.natural[,2] < 0.05) # natural, cold
table(p.natural[,3] < 0.05) # natural, gdd
table(p.artificial[,1] < 0.05) # artificial, annual
table(p.artificial[,2] < 0.05) # artificial, cold
table(p.artificial[,3] < 0.05) # artificial, gdd
```




* For bighead and silver carp, there were fewer data points (32 datapoints in total), but more subsample sets (10 sets of subsamples. This gave us 19 data points after subsampling with a much larger variation (due to a larger number of combinations). At extremes, we would have 13 artificial and 6 natural (if all subseting choose artificial); or 10 natural and 9 artificial (if all subsetting choose natural).

* So we use stratified sub-sampling to reduce this effect.



## Concluding points

1. Black carp: Using cold temperature have a better fit (higher R2). No preference over the four types of models. So we chose the simple linear model (Akaike weight = 51%).

2. Black carp: When separate the two conditions, we see a large increase in the R2 for the natural condition. The artificial condition alone did not have a significant relationship between log AAM and temperature.

3. Asian carp: The simple linear model is preferred (87% for annual, and 70% for cold). For grass carp, the simple linear model is preferred only for using the cold temperature (100%). For bs carp, there is no preference among the four models.

4. Asian carp: Using annual temperature is preferred (65% of times when annual temperature is the preferred). There is no preferrence for grass carp; but strong preference for bs carp using annual temperature.

5. Stratified sub-sampling reduced the large variation in AIC values in the bighead and silver carp combined dataset.
