---
title: "Spatial autocorrelation"
author: "Eddie Wu"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This .Rmd file is to test the spatial autocorrelation in our black carp data file, and try to find the best distance that reduces the spatial autocorrelation when doing subsampling. In this file, we:

1. Use moran's I test to test the global spatial autocorrelation in our dataset.

2. Use the correlog plots to examine how spatial autocorrelation changes with distances.

3. Use the variogram to further test point 2.

4. Subsample.



```{r library, warning=FALSE, message=FALSE}
library(gstat)
library(ggplot2)
library(dplyr)
library(spdep)
library(sp)
library(nlme)
library(ape)
library(MuMIn)
library(raster)
library(ncf)
library(knitr)
```

```{r data import}
## Import location data
location <- read.csv("location_no_temps.csv")
location <- unique(location) # remove duplicating locations

# Clean the data
carp <- read.csv("eddie_carp_new.csv")
carp.r <- carp %>% 
  filter(sex != "male") %>% # keep the non-male data points
  distinct(location, .keep_all = TRUE) # remove all repeating locations

## Download one file to get the spatial points
# (this defines what projection to use when converting to spatial objects)
tmin.1979 <- brick("cpc/tmin.1979.nc", varname = "tmin")
tmin.1979<-rotate(tmin.1979)
```


## Calculate the residuals from the linear model

```{r model}
# Define the model
lm.annual <- lm(log(carp.r$AAM)~carp.r$AnnualTemp) # Annual temperature
lm.cold <- lm(log(carp.r$AAM)~carp.r$ColdTemp) # Cold temperature
lm.warm <- lm(log(carp.r$AAM)~carp.r$WarmTemp) # Warm temperature

# See the results
summary(lm.annual)
summary(lm.cold)
summary(lm.warm)
```



## Global spatial autocorrelation

```{r global}
## Make spatial dataframe
coords <- data.frame("long"=location[,3],"lat"=location[,2])
df <- data.frame(a = 1:nrow(location[3]))
spatial.data <- SpatialPointsDataFrame(coords,df,proj4string = tmin.1979@crs)

# Get a distance matrix from all points
dists <- spDists(spatial.data, longlat = TRUE)

## Run the Moran.I test on the residuals
Moran.annual <- Moran.I(lm.annual$residuals, dists)
Moran.cold <- Moran.I(lm.cold$residuals, dists)

global.moran <- data.frame(
  Model = c("Moran.annual", "Moran.cold"),
  Observed = c(Moran.annual$observed, Moran.cold$observed),
  Expected = c(Moran.annual$expected, Moran.cold$expected),
  sd = c(Moran.annual$sd, Moran.cold$sd),
  p.value = c(Moran.annual$p.value, Moran.cold$p.value)
)

kable(global.moran)
```

There is no global spatial autocorrelation for the entire dataset.


## Correlog plots - local spatial autocorrelation

```{r correlog}
## Annual temperature
par(mfrow=c(2,1),mar=c(4,4,2,2))

# Use the coorelog function to develop the relationship
test <- correlog(coords$long, coords$lat, lm.annual$residuals,
                 increment=50, resamp=500, latlon=T)

# Plot with the entire distance range
plot(test, main="Annual Average Temperature Regression Residuals")
abline(h=0)
text(17400, min(test$correlation)+1, "A", cex=1.5)

# Reduce the distance range to 2500 km
plot(test, main="", xlim=c(0,2500))
abline(h=0)
text(2500, min(test$correlation)+1, "B", cex=1.5)


## Cold temperature
par(mfrow=c(2,1),mar=c(4,4,2,2))

# Use the coorelog function to develop the relationship
test <- correlog(coords$long, coords$lat, lm.cold$residuals,
                 increment=50, resamp=500, latlon=T)

# Plot with the entire distance range
plot(test, main="Cold Quarter Temperature Regression Residuals")
abline(h=0)
text(17400, min(test$correlation)+1, "A", cex=1.5)

# Reduce the distance range to 2500 km
plot(test, main="", xlim=c(0,2500))
abline(h=0)
text(2500, min(test$correlation)+1, "B", cex=1.5)
```

In general, the correlog plots suggest that for both temperature, 600 km is a good distance to reduce local spatial autocorrelation.


## Variogram

```{r variogram}
## Create a data frame to store the necessary information
vario.data <- data.frame(residuals = lm.annual$residuals,
                         x = coords$long, y = coords$lat)

# Turn them to spatial points using our projection
geo.spatial<-SpatialPoints(coords, proj4string = tmin.1979@crs)


## Variogram
vario <- variogram(vario.data$residuals~1, data = geo.spatial)
plot(vario)
```


If we pick 250 km as the cutoff distance for sub-sampling, we would have 20 location points.

If we pick 500 km as the cutoff distance for sub-sampling, we would have only 13 location points.


## Subsampling examination













