---
title: "TEMP temperature comparison"
author: "Eddie Wu"
date: '`r Sys.Date()`'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This .Rmd file is to show the results of water temperature comparison between the global model prediction and the tributary field observation for 11 Great Lake locations. (missing Saginaw data)

For all the graphs below, the black line is the mean weekly average temperature, the blue lines are the 95% confidence interval, the red dots are field observations by week.

```{r data import and setup}
library(ggplot2)
library(dplyr)
library(knitr)


tri.location <- read.csv("tributary locations.csv")
WT <- read.csv("water temperature clean/water_temperature_final_clean.csv",
               stringsAsFactors = TRUE)
```


## St. Louis River

``` {r st. louis}
# observed values
st.louis.field <- WT %>% filter(river == "stlouis")

# predicted values
st.louis.pred <- read.csv("water temperature clean/st_louis_model.csv")

## Plotting
ggplot(st.louis.pred, aes(x = weeks, y = temperature.avg))+
  geom_line(size = 1)+
  geom_line(aes(x = weeks, y = lower.CI), color = "blue", alpha = 0.5)+
  geom_line(aes(x = weeks, y = upper.CI), color = "blue", alpha = 0.5)+
  geom_point(st.louis.field, mapping = aes(x = weeks, y = temp), color = "red")+
  theme_bw()

## Calculate the Root Mean Square Error (RMSE)
RMSE.st.louis <- sqrt(mean((st.louis.field$temp - st.louis.pred$temperature.avg)^2))
RMSE.st.louis
```


## Fox River

``` {r fox}
# observed values
fox.field <- WT %>% filter(river == "fox")

# predicted values
fox.pred <- read.csv("water temperature clean/fox_model.csv")

## Plotting
ggplot(fox.pred, aes(x = weeks, y = temperature.avg))+
  geom_line(size = 1)+
  geom_line(aes(x = weeks, y = lower.CI), color = "blue", alpha = 0.5)+
  geom_line(aes(x = weeks, y = upper.CI), color = "blue", alpha = 0.5)+
  geom_point(fox.field, mapping = aes(x = weeks, y = temp), color = "red")+
  theme_bw()

## Calculate the Root Mean Square Error (RMSE)
RMSE.fox <- sqrt(mean((fox.field$temp - fox.pred$temperature.avg)^2))
RMSE.fox
```


## Portage-Burns Waterways

``` {r pb}
# observed values
pb.field <- WT %>% filter(river == "portage burns")

# predicted values
pb.pred <- read.csv("water temperature clean/pb_model.csv")

## Plotting
ggplot(pb.pred, aes(x = weeks, y = temperature.avg))+
  geom_line(size = 1)+
  geom_line(aes(x = weeks, y = lower.CI), color = "blue", alpha = 0.5)+
  geom_line(aes(x = weeks, y = upper.CI), color = "blue", alpha = 0.5)+
  geom_point(pb.field, mapping = aes(x = weeks, y = temp), color = "red")+
  theme_bw()

## Calculate the Root Mean Square Error (RMSE)
RMSE.pb <- sqrt(mean((pb.field$temp - pb.pred$temperature.avg)^2))
RMSE.pb
```


## Vermilion River

``` {r vermilion}
# observed values
vermilion.field <- WT %>% filter(river == "vermilion")

# predicted values
vermilion.pred <- read.csv("water temperature clean/vermilion_model.csv")

## Plotting
ggplot(vermilion.pred, aes(x = weeks, y = temperature.avg))+
  geom_line(size = 1)+
  geom_line(aes(x = weeks, y = lower.CI), color = "blue", alpha = 0.5)+
  geom_line(aes(x = weeks, y = upper.CI), color = "blue", alpha = 0.5)+
  geom_point(vermilion.field, mapping = aes(x = weeks, y = temp), color = "red")+
  theme_bw()

## Calculate the Root Mean Square Error (RMSE)
RMSE.vermilion <- 
  sqrt(mean((vermilion.field$temp - vermilion.pred$temperature.avg)^2))
RMSE.vermilion
```


## Genesee River

``` {r genesee}
# observed values
genesee.field <- WT %>% filter(river == "genesee")

# predicted values
genesee.pred <- read.csv("water temperature clean/genesee_model.csv")

## Plotting
ggplot(genesee.pred, aes(x = weeks, y = temperature.avg))+
  geom_line(size = 1)+
  geom_line(aes(x = weeks, y = lower.CI), color = "blue", alpha = 0.5)+
  geom_line(aes(x = weeks, y = upper.CI), color = "blue", alpha = 0.5)+
  geom_point(genesee.field, mapping = aes(x = weeks, y = temp), color = "red")+
  theme_bw()

## Calculate the Root Mean Square Error (RMSE)
RMSE.genesee <- sqrt(mean((genesee.field$temp - genesee.pred$temperature.avg)^2))
RMSE.genesee
```


## Big Creek River

``` {r big creek}
# observed values
big.creek.field <- WT %>% filter(river == "bigcreek")

# predicted values
big.creek.pred <- read.csv("water temperature clean/bigcreek_model.csv")

## Plotting
ggplot(big.creek.pred, aes(x = weeks, y = temperature.avg))+
  geom_line(size = 1)+
  geom_line(aes(x = weeks, y = lower.CI), color = "blue", alpha = 0.5)+
  geom_line(aes(x = weeks, y = upper.CI), color = "blue", alpha = 0.5)+
  geom_point(big.creek.field, mapping = aes(x = weeks, y = temp), color = "red")+
  theme_bw()

## Calculate the Root Mean Square Error (RMSE)
RMSE.big.creek <-
  sqrt(mean((big.creek.field$temp - big.creek.pred$temperature.avg)^2))
RMSE.big.creek
```


## Big Otter River

``` {r big otter}
# observed values
big.otter.field <- WT %>% filter(river == "bigotter")

# predicted values
big.otter.pred <- read.csv("water temperature clean/bigotter_model.csv")

## Plotting
ggplot(big.otter.pred, aes(x = weeks, y = temperature.avg))+
  geom_line(size = 1)+
  geom_line(aes(x = weeks, y = lower.CI), color = "blue", alpha = 0.5)+
  geom_line(aes(x = weeks, y = upper.CI), color = "blue", alpha = 0.5)+
  geom_point(big.otter.field, mapping = aes(x = weeks, y = temp), color = "red")+
  theme_bw()

## Calculate the Root Mean Square Error (RMSE)
RMSE.big.otter <-
  sqrt(mean((big.otter.field$temp - big.otter.pred$temperature.avg)^2))
RMSE.big.otter
```


## Still River

``` {r still}
# observed values
still.field <- WT %>% filter(river == "still")

# predicted values
still.pred <- read.csv("water temperature clean/still_model.csv")

## Plotting
ggplot(still.pred, aes(x = weeks, y = temperature.avg))+
  geom_line(size = 1)+
  geom_line(aes(x = weeks, y = lower.CI), color = "blue", alpha = 0.5)+
  geom_line(aes(x = weeks, y = upper.CI), color = "blue", alpha = 0.5)+
  geom_point(still.field, mapping = aes(x = weeks, y = temp), color = "red")+
  theme_bw()

## Calculate the Root Mean Square Error (RMSE)
RMSE.still <- sqrt(mean((still.field$temp - still.pred$temperature.avg)^2))
RMSE.still
```


## Mississagi River

``` {r mississagi}
# observed values
mississagi.field <- WT %>% filter(river == "mississagi")
str(mississagi.field)

# predicted values
mississagi.pred <- read.csv("water temperature clean/mississagi_model.csv")


## Plotting
ggplot(mississagi.pred, aes(x = weeks, y = temperature.avg))+
  geom_line(size = 1)+
  geom_line(aes(x = weeks, y = lower.CI), color = "blue", alpha = 0.5)+
  geom_line(aes(x = weeks, y = upper.CI), color = "blue", alpha = 0.5)+
  geom_point(mississagi.field, mapping = aes(x = weeks, y = temp), color = "red")+
  theme_bw()


## Calculate the Root Mean Square Error (RMSE)
RMSE.mississagi <-
  sqrt(mean((mississagi.field$temp - mississagi.pred$temperature.avg)^2))
RMSE.mississagi
```


## Nipigon River

``` {r nipigon}
# observed values
nipigon.field <- WT %>% filter(river == "nipigon")
str(nipigon.field)

# predicted values
nipigon.pred <- read.csv("water temperature clean/nipigon_model.csv")

## Plotting
ggplot(nipigon.pred, aes(x = weeks, y = temperature.avg))+
  geom_line(size = 1)+
  geom_line(aes(x = weeks, y = lower.CI), color = "blue", alpha = 0.5)+
  geom_line(aes(x = weeks, y = upper.CI), color = "blue", alpha = 0.5)+
  geom_point(nipigon.field, mapping = aes(x = weeks, y = temp), color = "red")+
  theme_bw()

## Calculate the Root Mean Square Error (RMSE)
RMSE.nipigon <- sqrt(mean((nipigon.field$temp - nipigon.pred$temperature.avg)^2))
RMSE.nipigon
```


## Humber River

``` {r humber}
# observed values
humber.field <- WT %>% filter(river == "humber")

# predicted values
humber.pred <- read.csv("water temperature clean/humber_model.csv")

## Plotting
ggplot(humber.pred, aes(x = weeks, y = temperature.avg))+
  geom_line(size = 1)+
  geom_line(aes(x = weeks, y = lower.CI), color = "blue", alpha = 0.5)+
  geom_line(aes(x = weeks, y = upper.CI), color = "blue", alpha = 0.5)+
  geom_point(humber.field, mapping = aes(x = weeks, y = temp), color = "red")+
  theme_bw()

## Calculate the Root Mean Square Error (RMSE)
RMSE.humber <- sqrt(mean((humber.field$temp - humber.pred$temperature.avg)^2))
RMSE.humber
```


## Summary of RMSE for all locations

``` {r summary}
tri.location <- tri.location[tri.location$tributary.name != "Saginaw River",]

RMSE.summary <- data.frame(
  tributary = tri.location$tributary.name,
  lake = tri.location$corresponding.lake,
  RMSE = c(RMSE.st.louis, RMSE.fox, RMSE.pb, RMSE.vermilion, RMSE.genesee,
           RMSE.big.creek, RMSE.big.otter, RMSE.still, RMSE.mississagi,
           RMSE.nipigon, RMSE.humber)
)
kable(RMSE.summary)
```


